---
title: "Manifestation of the climate change in Texas from 1950 until today"
author: |
  | Hans-Christian Lindner
  | Institute for Geoinformatics
  | Münster, Germany
  | h_lind04@uni-muenster.de
date: "16th of March 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE,  message=FALSE}
#Loading in all important libraries
library("readxl")
library("maps")
library("maptools")
library("mapdata")
library("xts")
library("plyr")
library("forecast")
library("kableExtra")
setwd("data")
```

```{r, echo=FALSE,  message=FALSE, warning=FALSE}
#functions for the temperature timeseries
createTempIndexTemp = function(rawData){
  rawDataIndex = rawData$DATE
  rawDataIndex.date = as.Date(rawDataIndex)
  return(rawDataIndex.date)
}

fillMissingDatesTemp = function(dates, dataFrame){
  dateRange = seq(min(dates), max(dates), by = 1)
  allDates = data.frame(dateRange)
  colnames(allDates) = c("dateRange" = "DATE")
  dataFrame = merge(dataFrame, allDates, all=TRUE)
  dataFrame = dataFrame[!duplicated(dataFrame$DATE),]
  return(dataFrame)
}

deleteNAsTemp = function(frame){
  if(sum(is.na(frame)) > 0){
    frame = na.approx(frame)
  } else {
    print("no NA's")
  }
}

tsTrendTemp = function(start, dataFrame){
  tempTS = ts(dataFrame$TMAX, start=start, frequency=365.25)
  tempTS = na.approx(tempTS)
  stlTempTS = stl(tempTS, s.window = "periodic")
  trendTempTS = stlTempTS$time.series[,2]
  return(trendTempTS)
}

forecastTemp = function(xtsTemp){
  arimaTemp = auto.arima(xtsTemp)
  fcastTemp = forecast(arimaTemp, h=100)
  return(fcastTemp)
}

statisticsTemp = function(xtsTemp){
  mean = mean(xtsTemp)
  sd = sd(xtsTemp)
  min = min(xtsTemp)
  max = max(xtsTemp)
  var = var(xtsTemp)
  median = median(xtsTemp)
  firstQuartile = quantile(xtsTemp, 0.25)
  thirdQuartile = quantile(xtsTemp, 0.75)
  statistics = data.frame(mean, sd, min, max, var, median, firstQuartile, thirdQuartile)
  return(statistics)
}
```

```{r, echo=FALSE, fig.show="hide",  message=FALSE, warning=FALSE}
#analysis of the temperature timeseries
setwd("data")
# Angleton, Texas in the South East near Houston
# Temperature timeseries for Angleton in the SouthEast of Texas
angletonTemp = read_xlsx("timeseries/Angleton, SE/Angleton, TMax.xlsx")
angletonTempIndex = createTempIndexTemp(angletonTemp)

#fill missing dates
angletonTemp = fillMissingDatesTemp(angletonTempIndex, angletonTemp)

#create xts object
angletonTempIndex = createTempIndexTemp(angletonTemp)
tmpMaxAngleton = xts(angletonTemp$TMAX, angletonTempIndex)

#delete NA's
tmpMaxAngleton = deleteNAsTemp(tmpMaxAngleton)

#plot
plot(tmpMaxAngleton)

#plot acf and pacf of the timeseries
acfAngletonTemp = acf(tmpMaxAngleton, lag.max = 365)
pacfAngletonTemp = pacf(tmpMaxAngleton)

#create trend
start = as.Date("1950-01-01") 
trendTmpAngleton = tsTrendTemp(start, angletonTemp)

#forecast
fcastTempAngleton = forecastTemp(tmpMaxAngleton)

#create statistics
statisticsAngletonTemp = statisticsTemp(tmpMaxAngleton)


# Benbrook Dam, Texas in the North East near Dallas
# Temperature timeseries for Benbrook in the NorthEast of Texas
BenbrookTemp = read_xlsx("timeseries/Benbrook, NE/Benbrook, TMax.xlsx")
BenbrookTempIndex = createTempIndexTemp(BenbrookTemp)


#fill missing dates
BenbrookTemp = fillMissingDatesTemp(BenbrookTempIndex, BenbrookTemp)

#create xts object
BenbrookTempIndex = createTempIndexTemp(BenbrookTemp)
tmpMaxBenbrook = xts(BenbrookTemp$TMAX, BenbrookTempIndex)

#delete NA's
tmpMaxBenbrook = deleteNAsTemp(tmpMaxBenbrook)

#plot
plot(tmpMaxBenbrook)

#plot acf and pacf of the timeseries
acfBenbrookTemp = acf(tmpMaxBenbrook, lag.max = 365)
pacfBenbrookTemp = pacf(tmpMaxBenbrook)

#create trend
start = as.Date("1950-07-03") 
trendTmpBenbrook = tsTrendTemp(start, BenbrookTemp)

#forecast
fcastTempBenbrook = forecastTemp(tmpMaxBenbrook)

#create statistics
statisticsBenbrookTemp = statisticsTemp(tmpMaxBenbrook)

# Fort Stockton, Texas in the North West near the border to New Mexico
# Temperature timeseries for Fort Stockton in the NorthWest of Texas
fortStocktonTemp = read_xlsx("timeseries/FortStockton, NW/FortStockton, TMax.xlsx")
fortStocktonTempIndex = createTempIndexTemp(fortStocktonTemp)


#fill missing dates
fortStocktonTemp = fillMissingDatesTemp(fortStocktonTempIndex, fortStocktonTemp)

#create xts object
fortStocktonTempIndex = createTempIndexTemp(fortStocktonTemp)
tmpMaxfortStockton = xts(fortStocktonTemp$TMAX, fortStocktonTempIndex)

#delete NA's
tmpMaxfortStockton = deleteNAsTemp(tmpMaxfortStockton)

#plot
plot(tmpMaxfortStockton)

#plot acf and pacf of the timeseries
acffortStocktonTemp = acf(tmpMaxfortStockton, lag.max = 365)
pacffortStocktonTemp = pacf(tmpMaxfortStockton)

#create trend
start = as.Date("1950-01-01") 
trendTmpfortStockton = tsTrendTemp(start, fortStocktonTemp)

#forecast
fcastTempfortStockton = forecastTemp(tmpMaxfortStockton)

#create statistics
statisticsfortStocktonTemp = statisticsTemp(tmpMaxfortStockton)

# Mccook, Texas in the South West near the coast of the Golf of Mexico and the Mexican border
# Temperature timeseries for Mccook in the South West of Texas
MccookTemp = read_xlsx("timeseries/Mccook, SW/Mccook, TMax.xlsx")
MccookTempIndex = createTempIndexTemp(MccookTemp)


#fill missing dates
MccookTemp = fillMissingDatesTemp(MccookTempIndex, MccookTemp)

#create xts object
MccookTempIndex = createTempIndexTemp(MccookTemp)
tmpMaxMccook = xts(MccookTemp$TMAX, MccookTempIndex)

#delete NA's
tmpMaxMccook = deleteNAsTemp(tmpMaxMccook)

#plot
plot(tmpMaxMccook)

#plot acf and pacf of the timeseries
acfMccookTemp = acf(tmpMaxMccook, lag.max = 365)
pacfMccookTemp = pacf(tmpMaxMccook)

#create trend
start = as.Date("1950-01-01") 
trendTmpMccook = tsTrendTemp(start, MccookTemp)

#forecast
fcastTempMccook = forecastTemp(tmpMaxMccook)

#create statistics
statisticsMccookTemp = statisticsTemp(tmpMaxMccook)

#boxplot of temperature timeseries
combineTmpTS = merge(tmpMaxAngleton, tmpMaxBenbrook, tmpMaxfortStockton, tmpMaxMccook)
combineTmpTS = rename(combineTmpTS, replace=c("tmpMaxAngleton" = "Angleton", "tmpMaxBenbrook" = "Benbrook", "tmpMaxfortStockton" = "Fort Stockton", "tmpMaxMccook" = "Mccook"))
boxplot(coredata(combineTmpTS))

#linear models of the time series
linTrendAngletonTemp = lm(trendTmpAngleton~time(trendTmpAngleton))
linTrendBenbrookTemp = lm(trendTmpBenbrook~time(trendTmpBenbrook))
linTrendfortStocktonTemp = lm(trendTmpfortStockton~time(trendTmpfortStockton))
linTrendMccookTemp = lm(trendTmpMccook~time(trendTmpMccook))

#plot trend of time series and compare linear regression with slope
par(mfrow=c(1,1))
plot(trendTmpAngleton)
abline(linTrendAngletonTemp, col="red")
slopeAngletonTemp = linTrendAngletonTemp$coefficients[2]
plot(trendTmpBenbrook)
abline(linTrendBenbrookTemp, col="red")
slopeBenbrookTemp = linTrendBenbrookTemp$coefficients[2]
plot(trendTmpfortStockton)
abline(linTrendfortStocktonTemp, col="red")
slopefortStocktonTemp = linTrendfortStocktonTemp$coefficients[2]
plot(trendTmpMccook)
abline(linTrendMccookTemp, col="red")
slopeMccookTemp = linTrendMccookTemp$coefficients[2]
slopeTemp = data.frame(slopeAngletonTemp, slopeBenbrookTemp, slopefortStocktonTemp, slopeMccookTemp)
```

```{r, echo=FALSE,  message=FALSE, warning=FALSE}
#functions for the precipitation timeseries
createPrcpIndexPrcp = function(rawData){
  rawDataIndex = rawData$DATE
  rawDataIndex.date = as.Date(rawDataIndex)
  return(rawDataIndex.date)
}

fillMissingDatesPrcp = function(dates, dataFrame){
  dateRange = seq(min(dates), max(dates), by = 1)
  allDates = data.frame(dateRange)
  colnames(allDates) = c("dateRange" = "DATE")
  dataFrame = merge(dataFrame, allDates, all=TRUE)
  dataFrame = dataFrame[!duplicated(dataFrame$DATE),]
  return(dataFrame)
}

deleteNAsPrcp = function(frame){
  if(sum(is.na(frame)) > 0){
    frame = na.approx(frame)
  } else {
    print("no NA's")
  }
}

tsTrendPrcp = function(start, dataFrame){
  PrcpTS = ts(dataFrame$PRCP, start=start, frequency=365.25)
  PrcpTS = na.approx(PrcpTS)
  stlPrcpTS = stl(PrcpTS, s.window = "periodic")
  trendPrcpTS = stlPrcpTS$time.series[,2]
  return(trendPrcpTS)
}

forecastPrcp = function(xtsPrcp){
  arimaPrcp = auto.arima(xtsPrcp)
  fcastPrcp = forecast(arimaPrcp, h=100)
  return(fcastPrcp)
}

statisticsPrcp = function(xtsPrcp){
  mean = mean(xtsPrcp)
  sd = sd(xtsPrcp)
  min = min(xtsPrcp)
  max = max(xtsPrcp)
  var = var(xtsPrcp)
  median = median(xtsPrcp)
  firstQuartile = quantile(xtsPrcp, 0.25)
  thirdQuartile = quantile(xtsPrcp, 0.75)
  statistics = data.frame(mean, sd, min, max, var, median, firstQuartile, thirdQuartile)
  return(statistics)
}
```

```{r, echo=FALSE, fig.show="hide",  message=FALSE, warning=FALSE}
#analysis of the precipitation timeseries
setwd("data")
# Angleton, Texas in the South East near Houston
# Precipitation timeseries for Angleton in the SouthEast of Texas
angletonPrcp = read_xlsx("timeseries/Angleton, SE/Angleton, Prcp.xlsx")
angletonPrcpIndex = createPrcpIndexPrcp(angletonPrcp)

#fill missing dates
angletonPrcp = fillMissingDatesPrcp(angletonPrcpIndex, angletonPrcp)

#create an xts object
angletonPrcpIndex = createPrcpIndexPrcp(angletonPrcp)
PrcpAngleton = xts(angletonPrcp$PRCP, angletonPrcpIndex)

#delete NA's
PrcpAngleton = deleteNAsPrcp(PrcpAngleton)

#plot
plot(PrcpAngleton)

#plot acf and pacf of the timeseries
acfAngletonPrcp = acf(PrcpAngleton, lag.max = 365)
pacfAngletonPrcp = pacf(PrcpAngleton)

#create trend
start = as.Date("1950-01-01")  
trendPrcpAngleton = tsTrendPrcp(start, angletonPrcp)

#forecast
fcastPrcpAngleton = forecastPrcp(PrcpAngleton)

#statistics
statisticsAngletonPrcp = statisticsPrcp(PrcpAngleton)


# Benbrook Dam, Texas in the North East near Dallas
# Precipitation timeseries for Benbrook in the NorthEast of Texas
BenbrookPrcp = read_xlsx("timeseries/Benbrook, NE/Benbrook, Prcp.xlsx")
BenbrookPrcpIndex = createPrcpIndexPrcp(BenbrookPrcp)

#fill missing dates
BenbrookPrcp = fillMissingDatesPrcp(BenbrookPrcpIndex, BenbrookPrcp)

#create an xts object
BenbrookPrcpIndex = createPrcpIndexPrcp(BenbrookPrcp)
PrcpBenbrook = xts(BenbrookPrcp$PRCP, BenbrookPrcpIndex)

#delete NA's
PrcpBenbrook = deleteNAsPrcp(PrcpBenbrook)

#plot
plot(PrcpBenbrook)

#plot acf and pacf of the timeseries
acfBenbrookPrcp = acf(PrcpBenbrook, lag.max = 365)
pacfBenbrookPrcp = pacf(PrcpBenbrook)

#create trend
start = as.Date("1950-01-01")  
trendPrcpBenbrook = tsTrendPrcp(start, BenbrookPrcp)

#forecast
fcastPrcpBenbrook = forecastPrcp(PrcpBenbrook)

#statistics
statisticsBenbrookPrcp = statisticsPrcp(PrcpBenbrook)



# Fort Stockton, Texas in the North West near the border to New Mexico
# Precipitation timeseries for Fort Stockton in the NorthWest of Texas
fortStocktonPrcp = read_xlsx("timeseries/fortStockton, NW/fortStockton, Prcp.xlsx")
fortStocktonPrcpIndex = createPrcpIndexPrcp(fortStocktonPrcp)

#fill missing dates
fortStocktonPrcp = fillMissingDatesPrcp(fortStocktonPrcpIndex, fortStocktonPrcp)

#create an xts object
fortStocktonPrcpIndex = createPrcpIndexPrcp(fortStocktonPrcp)
PrcpfortStockton = xts(fortStocktonPrcp$PRCP, fortStocktonPrcpIndex)

#delete NA's
PrcpfortStockton = deleteNAsPrcp(PrcpfortStockton)

#plot
plot(PrcpfortStockton)

#plot acf and pacf of the timeseries
acffortStocktonPrcp = acf(PrcpfortStockton, lag.max = 365)
pacffortStocktonPrcp = pacf(PrcpfortStockton)

#create trend
start = as.Date("1950-01-01")  
trendPrcpfortStockton = tsTrendPrcp(start, fortStocktonPrcp)

#forecast
fcastPrcpfortStockton = forecastPrcp(PrcpfortStockton)

#statistics
statisticsfortStocktonPrcp = statisticsPrcp(PrcpfortStockton)



# Mccook, Texas in the South West near the coast of the Golf of Mexico and the Mexican border
# Precipitation timeseries for Mccook in the South West of Texas
MccookPrcp = read_xlsx("timeseries/Mccook, SW/Mccook, Prcp.xlsx")
MccookPrcpIndex = createPrcpIndexPrcp(MccookPrcp)

#fill missing dates
MccookPrcp = fillMissingDatesPrcp(MccookPrcpIndex, MccookPrcp)

#create an xts object
MccookPrcpIndex = createPrcpIndexPrcp(MccookPrcp)
PrcpMccook = xts(MccookPrcp$PRCP, MccookPrcpIndex)

#delete NA's
PrcpMccook = deleteNAsPrcp(PrcpMccook)

#plot
plot(PrcpMccook)

#plot acf and pacf of the timeseries
acfMccookPrcp = acf(PrcpMccook, lag.max = 365)
pacfMccookPrcp = pacf(PrcpMccook)

#create trend
start = as.Date("1950-01-01")  
trendPrcpMccook = tsTrendPrcp(start, MccookPrcp)

#forecast
fcastPrcpMccook = forecastPrcp(PrcpMccook)

#statistics
statisticsMccookPrcp = statisticsPrcp(PrcpMccook)


#boxplot of precipitation timeseries
combinePrcpTS = merge(PrcpAngleton, PrcpBenbrook, PrcpfortStockton, PrcpMccook)
combinePrcpTS = rename(combinePrcpTS, replace=c("PrcpAngleton" = "Angleton", "PrcpBenbrook" = "Benbrook", "PrcpfortStockton" = "Fort Stockton", "PrcpMccook" = "Mccook"))
boxplot(coredata(combinePrcpTS))

#linear models of the time series
linTrendAngletonPrcp = lm(trendPrcpAngleton~time(trendPrcpAngleton))
linTrendBenbrookPrcp = lm(trendPrcpBenbrook~time(trendPrcpBenbrook))
linTrendfortStocktonPrcp = lm(trendPrcpfortStockton~time(trendPrcpfortStockton))
linTrendMccookPrcp = lm(trendPrcpMccook~time(trendPrcpMccook))

#plot trend of time series and compare linear regression with slope
par(mfrow=c(1,1))
plot(trendPrcpAngleton)
abline(linTrendAngletonPrcp, col="red")
slopeAngletonPrcp = linTrendAngletonPrcp$coefficients[2]
plot(trendPrcpBenbrook)
abline(linTrendBenbrookPrcp, col="red")
slopeBenbrookPrcp = linTrendBenbrookPrcp$coefficients[2]
plot(trendPrcpfortStockton)
abline(linTrendfortStocktonPrcp, col="red")
slopefortStocktonPrcp = linTrendfortStocktonPrcp$coefficients[2]
plot(trendPrcpMccook)
abline(linTrendMccookPrcp, col="red")
slopeMccookPrcp = linTrendMccookPrcp$coefficients[2]
slopePrcp = data.frame(slopeAngletonPrcp, slopeBenbrookPrcp, slopefortStocktonPrcp, slopeMccookPrcp)
```

```{r, echo=FALSE,  message=FALSE, warning=FALSE}
#functions for the evaporation timeseries
createEvapIndexEvap = function(rawData){
  rawDataIndex = rawData$DATE
  rawDataIndex.date = as.Date(rawDataIndex)
  return(rawDataIndex.date)
}

fillMissingDatesEvap = function(dates, dataFrame){
  dateRange = seq(min(dates), max(dates), by = 1)
  allDates = data.frame(dateRange)
  colnames(allDates) = c("dateRange" = "DATE")
  dataFrame = merge(dataFrame, allDates, all=TRUE)
  dataFrame = dataFrame[!duplicated(dataFrame$DATE),]
  return(dataFrame)
}

deleteNAsEvap = function(frame){
  if(sum(is.na(frame)) > 0){
    frame = na.approx(frame)
  } else {
    print("no NA's")
  }
}

tsTrendEvap = function(start, dataFrame){
  EvapTS = ts(dataFrame$EVAP, start=start, frequency=365.25)
  EvapTS = na.approx(EvapTS)
  stlEvapTS = stl(EvapTS, s.window = "periodic")
  trendEvapTS = stlEvapTS$time.series[,2]
  return(trendEvapTS)
}

forecastEvap = function(xtsEvap){
  arimaEvap = auto.arima(xtsEvap)
  fcastEvap = forecast(arimaEvap, h=100)
  return(fcastEvap)
}

statisticsEvap = function(xtsEvap){
  mean = mean(xtsEvap)
  sd = sd(xtsEvap)
  min = min(xtsEvap)
  max = max(xtsEvap)
  var = var(xtsEvap)
  median = median(xtsEvap)
  firstQuartile = quantile(xtsEvap, 0.25)
  thirdQuartile = quantile(xtsEvap, 0.75)
  statistics = data.frame(mean, sd, min, max, var, median, firstQuartile, thirdQuartile)
  return(statistics)
}
```

```{r, echo=FALSE, fig.show="hide",  message=FALSE, warning=FALSE}
#analysis of the evaporation timeseries
setwd("data")
# Angleton, Texas in the South East near Houston
# Evaporation timeseries for Angleton in the SouthEast of Texas
angletonEvap = read_xlsx("timeseries/Angleton, SE/Angleton, Evap.xlsx")
angletonEvapIndex = createEvapIndexEvap(angletonEvap)

#fill missing dates
angletonEvap = fillMissingDatesEvap(angletonEvapIndex, angletonEvap)

#create an xts object
angletonEvapIndex = createEvapIndexEvap(angletonEvap)
EvapAngleton = xts(angletonEvap$EVAP, angletonEvapIndex)

#delete NA's
EvapAngleton = deleteNAsEvap(EvapAngleton)

#plot
plot(EvapAngleton)

#plot acf and pacf of the timeseries
acfAngletonEvap = acf(EvapAngleton, lag.max = 365)
pacfAngletonEvap = pacf(EvapAngleton)

#create trend
start = as.Date("1953-01-01")
trendEvapAngleton = tsTrendEvap(start, angletonEvap)

#forecast
fcastEvapAngleton = forecastEvap(EvapAngleton)

#statistics
statisticsAngletonEvap = statisticsEvap(EvapAngleton)


# Benbrook Dam, Texas in the North East near Dallas
# Evaporation timeseries for Benbrook in the NorthEast of Texas
BenbrookEvap = read_xlsx("timeseries/Benbrook, NE/Benbrook, Evap.xlsx")
BenbrookEvapIndex = createEvapIndexEvap(BenbrookEvap)

#fill missing dates
BenbrookEvap = fillMissingDatesEvap(BenbrookEvapIndex, BenbrookEvap)

#create an xts object
BenbrookEvapIndex = createEvapIndexEvap(BenbrookEvap)
EvapBenbrook = xts(BenbrookEvap$EVAP, BenbrookEvapIndex)

#delete NA's
EvapBenbrook = deleteNAsEvap(EvapBenbrook)

#plot
plot(EvapBenbrook)

#plot acf and pacf of the timeseries
acfBenbrookEvap = acf(EvapBenbrook, lag.max = 365)
pacfBenbrookEvap = pacf(EvapBenbrook)

#create trend
start = as.Date("1953-01-01")
trendEvapBenbrook = tsTrendEvap(start, BenbrookEvap)

#forecast
fcastEvapBenbrook = forecastEvap(EvapBenbrook)

#statistics
statisticsBenbrookEvap = statisticsEvap(EvapBenbrook)



# Fort Stockton, Texas in the North West near the border to New Mexico
# Evaporation timeseries for Fort Stockton in the NorthWest of Texas
fortStocktonEvap = read_xlsx("timeseries/fortStockton, NW/fortStockton, Evap.xlsx")
fortStocktonEvapIndex = createEvapIndexEvap(fortStocktonEvap)

#fill missing dates
fortStocktonEvap = fillMissingDatesEvap(fortStocktonEvapIndex, fortStocktonEvap)

#create an xts object
fortStocktonEvapIndex = createEvapIndexEvap(fortStocktonEvap)
EvapfortStockton = xts(fortStocktonEvap$EVAP, fortStocktonEvapIndex)

#delete NA's
EvapfortStockton = deleteNAsEvap(EvapfortStockton)

#plot
plot(EvapfortStockton)

#plot acf and pacf of the timeseries
acffortStocktonEvap = acf(EvapfortStockton, lag.max = 365)
pacffortStocktonEvap = pacf(EvapfortStockton)

#create trend
start = as.Date("1953-01-01")
trendEvapfortStockton = tsTrendEvap(start, fortStocktonEvap)

#forecast
fcastEvapfortStockton = forecastEvap(EvapfortStockton)

#statistics
statisticsfortStocktonEvap = statisticsEvap(EvapfortStockton)



# Mccook, Texas in the South West near the coast of the Golf of Mexico and the Mexican border
# Evaporation timeseries for Mccook in the South West of Texas
MccookEvap = read_xlsx("timeseries/Mccook, SW/Mccook, Evap.xlsx")
MccookEvapIndex = createEvapIndexEvap(MccookEvap)

#fill missing dates
MccookEvap = fillMissingDatesEvap(MccookEvapIndex, MccookEvap)

#create an xts object
MccookEvapIndex = createEvapIndexEvap(MccookEvap)
EvapMccook = xts(MccookEvap$EVAP, MccookEvapIndex)

#delete NA's
EvapMccook = deleteNAsEvap(EvapMccook)

#plot
plot(EvapMccook)

#plot acf and pacf of the timeseries
acfMccookEvap = acf(EvapMccook, lag.max = 365)
pacfMccookEvap = pacf(EvapMccook)

#create trend
start = as.Date("1953-01-01")
trendEvapMccook = tsTrendEvap(start, MccookEvap)

#forecast
fcastEvapMccook = forecastEvap(EvapMccook)

#statistics
statisticsMccookEvap = statisticsEvap(EvapMccook)


#boxplot of evaporation timeseries
combineEvapTS = merge(EvapAngleton, EvapBenbrook, EvapfortStockton, EvapMccook)
combineEvapTS = rename(combineEvapTS, replace=c("EvapAngleton" = "Angleton", "EvapBenbrook" = "Benbrook", "EvapfortStockton" = "Fort Stockton", "EvapMccook" = "Mccook"))
boxplot(coredata(combineEvapTS))

#linear models of the time series
linTrendAngletonEvap = lm(trendEvapAngleton~time(trendEvapAngleton))
linTrendBenbrookEvap = lm(trendEvapBenbrook~time(trendEvapBenbrook))
linTrendfortStocktonEvap = lm(trendEvapfortStockton~time(trendEvapfortStockton))
linTrendMccookEvap = lm(trendEvapMccook~time(trendEvapMccook))

#plot trend of time series and compare linear regression with slope
par(mfrow=c(1,1))
plot(trendEvapAngleton)
abline(linTrendAngletonEvap, col="red")
slopeAngletonEvap = linTrendAngletonEvap$coefficients[2]
plot(trendEvapBenbrook)
abline(linTrendBenbrookEvap, col="red")
slopeBenbrookEvap = linTrendBenbrookEvap$coefficients[2]
plot(trendEvapfortStockton)
abline(linTrendfortStocktonEvap, col="red")
slopefortStocktonEvap = linTrendfortStocktonEvap$coefficients[2]
plot(trendEvapMccook)
abline(linTrendMccookEvap, col="red")
slopeMccookEvap = linTrendMccookEvap$coefficients[2]
slopeEvap = data.frame(slopeAngletonEvap, slopeBenbrookEvap, slopefortStocktonEvap, slopeMccookEvap)
```

```{r, echo=FALSE,  message=FALSE, warning=FALSE}
#functions for the wind movement timeseries
createWindIndexWind = function(rawData){
  rawDataIndex = rawData$DATE
  rawDataIndex.date = as.Date(rawDataIndex)
  return(rawDataIndex.date)
}

fillMissingDatesWind = function(dates, dataFrame){
  dateRange = seq(min(dates), max(dates), by = 1)
  allDates = data.frame(dateRange)
  colnames(allDates) = c("dateRange" = "DATE")
  dataFrame = merge(dataFrame, allDates, all=TRUE)
  dataFrame = dataFrame[!duplicated(dataFrame$DATE),]
  return(dataFrame)
}

deleteNAsWind = function(frame){
  if(sum(is.na(frame)) > 0){
    frame = na.approx(frame)
  } else {
    print("no NA's")
  }
}

tsTrendWind = function(start, dataFrame){
  WindTS = ts(dataFrame$WDMV, start=start, frequency=365.25)
  WindTS = na.approx(WindTS)
  stlWindTS = stl(WindTS, s.window = "periodic")
  trendWindTS = stlWindTS$time.series[,2]
  return(trendWindTS)
}

forecastWind = function(xtsWind){
  arimaWind = auto.arima(xtsWind)
  fcastWind = forecast(arimaWind, h=100)
  return(fcastWind)
}

statisticsWind = function(xtsWind){
  mean = mean(xtsWind)
  sd = sd(xtsWind)
  min = min(xtsWind)
  max = max(xtsWind)
  var = var(xtsWind)
  median = median(xtsWind)
  firstQuartile = quantile(xtsWind, 0.25)
  thirdQuartile = quantile(xtsWind, 0.75)
  statistics = data.frame(mean, sd, min, max, var, median, firstQuartile, thirdQuartile)
  return(statistics)
}
```

```{r, echo=FALSE, fig.show="hide",  message=FALSE, warning=FALSE}
setwd("data")
#analysis of the wind movement timeseries
# Angleton, Texas in the South East near Houston
# Wind timeseries for Angleton in the SouthEast of Texas
angletonWind = read_xlsx("timeseries/Angleton, SE/Angleton, Wind.xlsx")
angletonWindIndex = createWindIndexWind(angletonWind)

#fill missing dates
angletonWind = fillMissingDatesWind(angletonWindIndex, angletonWind)

#create an xts object
angletonWindIndex = createWindIndexWind(angletonWind)
WindAngleton = xts(angletonWind$WDMV, angletonWindIndex)

#delete NA's
WindAngleton = deleteNAsWind(WindAngleton)

#plot
plot(WindAngleton)

#plot acf and pacf of the timeseries
acfAngletonWind = acf(WindAngleton, lag.max = 365)
pacfAngletonWind = pacf(WindAngleton)

#create trend
start = as.Date("1953-01-01") 
trendWindAngleton = tsTrendWind(start, angletonWind)

#forecast
fcastWindAngleton = forecastWind(WindAngleton)

#statistics
statisticsAngletonWind = statisticsWind(WindAngleton)




# Benbrook Dam, Texas in the North East near Dallas
# Precipitation timeseries for Benbrook in the NorthEast of Texas
BenbrookWind = read_xlsx("timeseries/Benbrook, NE/Benbrook, Wind.xlsx")
BenbrookWindIndex = createWindIndexWind(BenbrookWind)

#fill missing dates
BenbrookWind = fillMissingDatesWind(BenbrookWindIndex, BenbrookWind)

#create an xts object
BenbrookWindIndex = createWindIndexWind(BenbrookWind)
WindBenbrook = xts(BenbrookWind$WDMV, BenbrookWindIndex)

#delete NA's
WindBenbrook = deleteNAsWind(WindBenbrook)

#plot
plot(WindBenbrook)

#plot acf and pacf of the timeseries
acfBenbrookWind = acf(WindBenbrook, lag.max = 365)
pacfBenbrookWind = pacf(WindBenbrook)

#create trend
start = as.Date("1979-01-01") 
trendWindBenbrook = tsTrendWind(start, BenbrookWind)

#forecast
fcastWindBenbrook = forecastWind(WindBenbrook)

#statistics
statisticsBenbrookWind = statisticsWind(WindBenbrook)



# Fort Stockton, Texas in the North West near the border to New Mexico
# Precipitation timeseries for Fort Stockton in the NorthWest of Texas
fortStocktonWind = read_xlsx("timeseries/fortStockton, NW/fortStockton, Wind.xlsx")
fortStocktonWindIndex = createWindIndexWind(fortStocktonWind)

#fill missing dates
fortStocktonWind = fillMissingDatesWind(fortStocktonWindIndex, fortStocktonWind)

#create an xts object
fortStocktonWindIndex = createWindIndexWind(fortStocktonWind)
WindfortStockton = xts(fortStocktonWind$WDMV, fortStocktonWindIndex)

#delete NA's
WindfortStockton = deleteNAsWind(WindfortStockton)

#plot
plot(WindfortStockton)

#plot acf and pacf of the timeseries
acffortStocktonWind = acf(WindfortStockton, lag.max = 365)
pacffortStocktonWind = pacf(WindfortStockton)

#create trend
start = as.Date("1950-01-01")
trendWindfortStockton = tsTrendWind(start, fortStocktonWind)

#forecast
fcastWindfortStockton = forecastWind(WindfortStockton)

#statistics
statisticsfortStocktonWind = statisticsWind(WindfortStockton)



# Mccook, Texas in the South West near the coast of the Golf of Mexico and the Mexican border
# Precipitation timeseries for Mccook in the South West of Texas
MccookWind = read_xlsx("timeseries/Mccook, SW/Mccook, Wind.xlsx")
MccookWindIndex = createWindIndexWind(MccookWind)

#fill missing dates
MccookWind = fillMissingDatesWind(MccookWindIndex, MccookWind)

#create an xts object
MccookWindIndex = createWindIndexWind(MccookWind)
WindMccook = xts(MccookWind$WDMV, MccookWindIndex)

#delete NA's
WindMccook = deleteNAsWind(WindMccook)

#plot
plot(WindMccook)

#plot acf and pacf of the timeseries
acfMccookWind = acf(WindMccook, lag.max = 365)
pacfMccookWind = pacf(WindMccook)

#create trend
start = as.Date("1963-01-01")
trendWindMccook = tsTrendWind(start, MccookWind)

#forecast
fcastWindMccook = forecastWind(WindMccook)

#statistics
statisticsMccookWind = statisticsWind(WindMccook)

#boxplot of precipitation timeseries
combineWindTS = merge(WindAngleton, WindBenbrook, WindfortStockton, WindMccook)
combineWindTS = rename(combineWindTS, replace=c("WindAngleton" = "Angleton", "WindBenbrook" = "Benbrook", "WindfortStockton" = "Fort Stockton", "WindMccook" = "Mccook"))
boxplot(coredata(combineWindTS))

#linear models of the time series
linTrendAngletonWind = lm(trendWindAngleton~time(trendWindAngleton))
linTrendBenbrookWind = lm(trendWindBenbrook~time(trendWindBenbrook))
linTrendfortStocktonWind = lm(trendWindfortStockton~time(trendWindfortStockton))
linTrendMccookWind = lm(trendWindMccook~time(trendWindMccook))

#plot trend of time series and compare linear regression with slope
par(mfrow=c(1,1))
plot(trendWindAngleton)
abline(linTrendAngletonWind, col="red")
slopeAngletonWind = linTrendAngletonWind$coefficients[2]
plot(trendWindBenbrook)
abline(linTrendBenbrookWind, col="red")
slopeBenbrookWind = linTrendBenbrookWind$coefficients[2]
plot(trendWindfortStockton)
abline(linTrendfortStocktonWind, col="red")
slopefortStocktonWind = linTrendfortStocktonWind$coefficients[2]
plot(trendWindMccook)
abline(linTrendMccookWind, col="red")
slopeMccookWind = linTrendMccookWind$coefficients[2]
slopeWind = data.frame(slopeAngletonWind, slopeBenbrookWind, slopefortStocktonWind, slopeMccookWind)
```

```{r, echo=FALSE, fig.show="hide",  message=FALSE, warning=FALSE}
#packages needed for the kriging part
library("gstat")
library("sp")
library("rgdal")
library("dplyr")

#creation of the grid for the kriging interpolation
texas = map('state', 'texas', fill=TRUE)
texas2 = data.frame(LONGITUDE = texas$x, LATITUDE = texas$y)
texas2 = na.omit(texas2)
coordinates(texas2) = texas2[c(1,2)]
proj4string(texas2) = CRS("+init=epsg:4326")
texas2 = spTransform(texas2, CRS("+proj=utm +zone=13 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))

texasGrid = spsample(texas2, type="regular")
tx = as.data.frame(coordinates(texasGrid))
names(tx) = c("LONGITUDE", "LATITUDE")
coordinates(tx) = tx[c(1,2)]
texasGrid = tx
proj4string(texasGrid) = CRS("+proj=utm +zone=13 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")

texasMap = map2SpatialPolygons(texas, ID=texas$names)
proj4string(texasMap) = CRS("+init=epsg:4326")
texasMap = spTransform(texasMap, CRS("+proj=utm +zone=13 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))

#function to compute basic statistics for data sets
statisticsCompute = function(dataSet){
  mean = mean(dataSet)
  sd = sd(dataSet)
  min = min(dataSet)
  max = max(dataSet)
  var = var(dataSet)
  median = median(dataSet)
  firstQuartile = quantile(dataSet, 0.25)
  thirdQuartile = quantile(dataSet, 0.75)
  statistics = data.frame(mean, sd, min, max, var, median, firstQuartile, thirdQuartile)
  return(statistics)
}
```

```{r, echo=FALSE,  message=FALSE, warning=FALSE}
#functions for the kriging interpolation of the temperature data
computeTempAverage = function(dataFrame){
  dataFrameSubset = mutate(dataFrame, month = format(DATE, "%m"))
  dataFrameMean = group_by(dataFrameSubset, STATION, NAME, LATITUDE, LONGITUDE, ELEVATION, month) %>% summarise(meanTemp = mean(TMAX, na.rm = T))
  dataFrameMean$meanTemp = round(dataFrameMean$meanTemp, digits=4)
  return(dataFrameMean)
}

coordinateTransformTemp = function(dataSet){
  coordinates(dataSet) = dataSet[c(4,3)]
  proj4string(dataSet) = CRS("+init=epsg:4326")
  dataSet = spTransform(dataSet, CRS("+proj=utm +zone=13 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))
  return(dataSet)
}

readDataTemp = function(path){
  rawData = read_xlsx(path)
  rawData$LATITUDE = as.numeric(rawData$LATITUDE)
  rawData$LONGITUDE = as.numeric(rawData$LONGITUDE)
  return(rawData)
}
```

```{r, echo=FALSE, fig.show="hide",  message=FALSE, warning=FALSE, results="hide"}
#analysis and kriging interpolation of the temperature data
setwd("data")
                                # JANUARY 2015 #

#Temperature values for January 2015 in Texas
JanTmpMax2015 = readDataTemp("2015/Jan/2015-Jan-TMax.xlsx")

#data frame with data for one day
JanTmpMax2015Subset = computeTempAverage(JanTmpMax2015)
JanTmpMax2015Subset = coordinateTransformTemp(JanTmpMax2015Subset)

#statistics
JanTmpMax2015Stats = statisticsCompute(JanTmpMax2015Subset$meanTemp)

# variogram for the one day
vJanTmpMax2015 = variogram(meanTemp ~ LATITUDE+LONGITUDE, JanTmpMax2015Subset)
plot(vJanTmpMax2015)

#fitting variogram
vJanTmpMax2015.fit = fit.variogram(vJanTmpMax2015, vgm(2, 'Lin', 0, 1))
plot(vJanTmpMax2015 ,vJanTmpMax2015.fit)

#kriging
JanTmpMax2015krige1 = krige(meanTemp ~ 1, JanTmpMax2015Subset, texasGrid, vJanTmpMax2015.fit)
spplot(JanTmpMax2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





                            # AUGUST 2015 #



#Temperature values for Auguary 2015 in Texas
AugTmpMax2015 = readDataTemp("2015/Aug/2015-Aug-TMax.xlsx")

#data frame with data for one day
AugTmpMax2015Subset = computeTempAverage(AugTmpMax2015)
AugTmpMax2015Subset = coordinateTransformTemp(AugTmpMax2015Subset)

#statistics
AugTmpMax2015Stats = statisticsCompute(AugTmpMax2015Subset$meanTemp)

# variogram for the one day
vAugTmpMax2015 = variogram(meanTemp ~ LATITUDE+LONGITUDE, AugTmpMax2015Subset)
plot(vAugTmpMax2015)

#fitting variogram
vAugTmpMax2015.fit = fit.variogram(vAugTmpMax2015, vgm(2, 'Lin', 0, 1))
plot(vAugTmpMax2015 ,vAugTmpMax2015.fit)

#kriging
AugTmpMax2015krige1 = krige(meanTemp ~ 1, AugTmpMax2015Subset, texasGrid, vAugTmpMax2015.fit)
spplot(AugTmpMax2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))




                                  # JANUARY 2005 #


#Temperature values for January 2005 in Texas
JanTmpMax2005 = readDataTemp("2005/Jan/2005-Jan-TMax.xlsx")

#data frame with data for one day
JanTmpMax2005Subset = computeTempAverage(JanTmpMax2005)
JanTmpMax2005Subset = coordinateTransformTemp(JanTmpMax2005Subset)

#statistics
JanTmpMax2005Stats = statisticsCompute(JanTmpMax2005Subset$meanTemp)

# variogram for the one day
vJanTmpMax2005 = variogram(meanTemp ~ LATITUDE+LONGITUDE, JanTmpMax2005Subset)
plot(vJanTmpMax2005)

#fitting variogram
vJanTmpMax2005.fit = fit.variogram(vJanTmpMax2005, vgm(2, 'Lin', 0, 1))
plot(vJanTmpMax2005 ,vJanTmpMax2005.fit)

#kriging
JanTmpMax2005krige1 = krige(meanTemp ~ 1, JanTmpMax2005Subset, texasGrid, vJanTmpMax2005.fit)
spplot(JanTmpMax2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





# AUGUST 2005 #



#Temperature values for Auguary 2005 in Texas
AugTmpMax2005 = readDataTemp("2005/Aug/2005-Aug-TMax.xlsx")

#data frame with data for one day
AugTmpMax2005Subset = computeTempAverage(AugTmpMax2005)
AugTmpMax2005Subset = coordinateTransformTemp(AugTmpMax2005Subset)

#statistics
AugTmpMax2005Stats = statisticsCompute(AugTmpMax2005Subset$meanTemp)

# variogram for the one day
vAugTmpMax2005 = variogram(meanTemp ~ LATITUDE+LONGITUDE, AugTmpMax2005Subset)
plot(vAugTmpMax2005)

#fitting variogram
vAugTmpMax2005.fit = fit.variogram(vAugTmpMax2005, vgm(2, 'Lin', 0, 1))
plot(vAugTmpMax2005 ,vAugTmpMax2005.fit)

#kriging
AugTmpMax2005krige1 = krige(meanTemp ~ 1, AugTmpMax2005Subset, texasGrid, vAugTmpMax2005.fit)
spplot(AugTmpMax2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





                                              # JANUARY 1995 #


#Temperature values for January 1995 in Texas
JanTmpMax1995 = readDataTemp("1995/Jan/1995-Jan-TMax.xlsx")

#data frame with data for one day
JanTmpMax1995Subset = computeTempAverage(JanTmpMax1995)
JanTmpMax1995Subset = coordinateTransformTemp(JanTmpMax1995Subset)

#statistics
JanTmpMax1995Stats = statisticsCompute(JanTmpMax1995Subset$meanTemp)

# variogram for the one day
vJanTmpMax1995 = variogram(meanTemp ~ LATITUDE+LONGITUDE, JanTmpMax1995Subset)
plot(vJanTmpMax1995)

#fitting variogram
vJanTmpMax1995.fit = fit.variogram(vJanTmpMax1995, vgm(2, 'Lin', 0, 1))
plot(vJanTmpMax1995 ,vJanTmpMax1995.fit)

#kriging
JanTmpMax1995krige1 = krige(meanTemp ~ 1, JanTmpMax1995Subset, texasGrid, vJanTmpMax1995.fit)
spplot(JanTmpMax1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





# AUGUST 1995 #



#Temperature values for Auguary 1995 in Texas
AugTmpMax1995 = readDataTemp("1995/Aug/1995-Aug-TMax.xlsx")

#data frame with data for one day
AugTmpMax1995Subset = computeTempAverage(AugTmpMax1995)
AugTmpMax1995Subset = coordinateTransformTemp(AugTmpMax1995Subset)

#statistics
AugTmpMax1995Stats = statisticsCompute(AugTmpMax1995Subset$meanTemp)

# variogram for the one day
vAugTmpMax1995 = variogram(meanTemp ~ LATITUDE+LONGITUDE, AugTmpMax1995Subset)
plot(vAugTmpMax1995)

#fitting variogram
vAugTmpMax1995.fit = fit.variogram(vAugTmpMax1995, vgm(2, 'Lin', 0, 1))
plot(vAugTmpMax1995 ,vAugTmpMax1995.fit)

#kriging
AugTmpMax1995krige1 = krige(meanTemp ~ 1, AugTmpMax1995Subset, texasGrid, vAugTmpMax1995.fit)
spplot(AugTmpMax1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





                                              # JANUARY 1985 #


#Temperature values for January 1985 in Texas
JanTmpMax1985 = readDataTemp("1985/Jan/1985-Jan-TMax.xlsx")

#data frame with data for one day
JanTmpMax1985Subset = computeTempAverage(JanTmpMax1985)
JanTmpMax1985Subset = coordinateTransformTemp(JanTmpMax1985Subset)

#statistics
JanTmpMax1985Stats = statisticsCompute(JanTmpMax1985Subset$meanTemp)

# variogram for the one day
vJanTmpMax1985 = variogram(meanTemp ~ LATITUDE+LONGITUDE, JanTmpMax1985Subset)
plot(vJanTmpMax1985)

#fitting variogram
vJanTmpMax1985.fit = fit.variogram(vJanTmpMax1985, vgm(2, 'Lin', 0, 1))
plot(vJanTmpMax1985 ,vJanTmpMax1985.fit)

#kriging
JanTmpMax1985krige1 = krige(meanTemp ~ 1, JanTmpMax1985Subset, texasGrid, vJanTmpMax1985.fit)
spplot(JanTmpMax1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





# AUGUST 1985 #



#Temperature values for Auguary 1985 in Texas
AugTmpMax1985 = readDataTemp("1985/Aug/1985-Aug-TMax.xlsx")

#data frame with data for one day
AugTmpMax1985Subset = computeTempAverage(AugTmpMax1985)
AugTmpMax1985Subset = coordinateTransformTemp(AugTmpMax1985Subset)

#statistics
AugTmpMax1985Stats = statisticsCompute(AugTmpMax1985Subset$meanTemp)

# variogram for the one day
vAugTmpMax1985 = variogram(meanTemp ~ LATITUDE+LONGITUDE, AugTmpMax1985Subset)
plot(vAugTmpMax1985)

#fitting variogram
vAugTmpMax1985.fit = fit.variogram(vAugTmpMax1985, vgm(2, 'Lin', 0, 1))
plot(vAugTmpMax1985 ,vAugTmpMax1985.fit)

#kriging
AugTmpMax1985krige1 = krige(meanTemp ~ 1, AugTmpMax1985Subset, texasGrid, vAugTmpMax1985.fit)
spplot(AugTmpMax1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





                                        # JANUARY 1975 #


#Temperature values for January 1975 in Texas
JanTmpMax1975 = readDataTemp("1975/Jan/1975-Jan-TMax.xlsx")

#data frame with data for one day
JanTmpMax1975Subset = computeTempAverage(JanTmpMax1975)
JanTmpMax1975Subset = coordinateTransformTemp(JanTmpMax1975Subset)

#statistics
JanTmpMax1975Stats = statisticsCompute(JanTmpMax1975Subset$meanTemp)

# variogram for the one day
vJanTmpMax1975 = variogram(meanTemp ~ LATITUDE+LONGITUDE, JanTmpMax1975Subset)
plot(vJanTmpMax1975)

#fitting variogram
vJanTmpMax1975.fit = fit.variogram(vJanTmpMax1975, vgm(2, 'Lin', 0, 1))
plot(vJanTmpMax1975 ,vJanTmpMax1975.fit)

#kriging
JanTmpMax1975krige1 = krige(meanTemp ~ 1, JanTmpMax1975Subset, texasGrid, vJanTmpMax1975.fit)
spplot(JanTmpMax1975krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





# AUGUST 1975 #



#Temperature values for Auguary 1975 in Texas
AugTmpMax1975 = readDataTemp("1975/Aug/1975-Aug-TMax.xlsx")

#data frame with data for one day
AugTmpMax1975Subset = computeTempAverage(AugTmpMax1975)
AugTmpMax1975Subset = coordinateTransformTemp(AugTmpMax1975Subset)

#statistics
AugTmpMax1975Stats = statisticsCompute(AugTmpMax1975Subset$meanTemp)

# variogram for the one day
vAugTmpMax1975 = variogram(meanTemp ~ LATITUDE+LONGITUDE, AugTmpMax1975Subset)
plot(vAugTmpMax1975)

#fitting variogram
vAugTmpMax1975.fit = fit.variogram(vAugTmpMax1975, vgm(2, 'Lin', 0, 1))
plot(vAugTmpMax1975 ,vAugTmpMax1975.fit)

#kriging
AugTmpMax1975krige1 = krige(meanTemp ~ 1, AugTmpMax1975Subset, texasGrid, vAugTmpMax1975.fit)
spplot(AugTmpMax1975krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





                                      # JANUARY 1965 #


#Temperature values for January 1965 in Texas
JanTmpMax1965 = readDataTemp("1965/Jan/1965-Jan-TMax.xlsx")

#data frame with data for one day
JanTmpMax1965Subset = computeTempAverage(JanTmpMax1965)
JanTmpMax1965Subset = coordinateTransformTemp(JanTmpMax1965Subset)

#statistics
JanTmpMax1965Stats = statisticsCompute(JanTmpMax1965Subset$meanTemp)

# variogram for the one day
vJanTmpMax1965 = variogram(meanTemp ~ LATITUDE+LONGITUDE, JanTmpMax1965Subset)
plot(vJanTmpMax1965)

#fitting variogram
vJanTmpMax1965.fit = fit.variogram(vJanTmpMax1965, vgm(2, 'Lin', 0, 1))
plot(vJanTmpMax1965 ,vJanTmpMax1965.fit)

#kriging
JanTmpMax1965krige1 = krige(meanTemp ~ 1, JanTmpMax1965Subset, texasGrid, vJanTmpMax1965.fit)
spplot(JanTmpMax1965krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





# AUGUST 1965 #



#Temperature values for Auguary 1965 in Texas
AugTmpMax1965 = readDataTemp("1965/Aug/1965-Aug-TMax.xlsx")

#data frame with data for one day
AugTmpMax1965Subset = computeTempAverage(AugTmpMax1965)
AugTmpMax1965Subset = coordinateTransformTemp(AugTmpMax1965Subset)

#statistics
AugTmpMax1965Stats = statisticsCompute(AugTmpMax1965Subset$meanTemp)

# variogram for the one day
vAugTmpMax1965 = variogram(meanTemp ~ LATITUDE+LONGITUDE, AugTmpMax1965Subset)
plot(vAugTmpMax1965)

#fitting variogram
vAugTmpMax1965.fit = fit.variogram(vAugTmpMax1965, vgm(2, 'Lin', 0, 1))
plot(vAugTmpMax1965 ,vAugTmpMax1965.fit)

#kriging
AugTmpMax1965krige1 = krige(meanTemp ~ 1, AugTmpMax1965Subset, texasGrid, vAugTmpMax1965.fit)
spplot(AugTmpMax1965krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





                                    # JANUARY 1955 #


#Temperature values for January 1955 in Texas
JanTmpMax1955 = readDataTemp("1955/Jan/1955-Jan-TMax.xlsx")

#data frame with data for one day
JanTmpMax1955Subset = computeTempAverage(JanTmpMax1955)
JanTmpMax1955Subset = coordinateTransformTemp(JanTmpMax1955Subset)

#statistics
JanTmpMax1955Stats = statisticsCompute(JanTmpMax1955Subset$meanTemp)

# variogram for the one day
vJanTmpMax1955 = variogram(meanTemp ~ LATITUDE+LONGITUDE, JanTmpMax1955Subset)
plot(vJanTmpMax1955)

#fitting variogram
vJanTmpMax1955.fit = fit.variogram(vJanTmpMax1955, vgm(2, 'Lin', 0, 1))
plot(vJanTmpMax1955 ,vJanTmpMax1955.fit)

#kriging
JanTmpMax1955krige1 = krige(meanTemp ~ 1, JanTmpMax1955Subset, texasGrid, vJanTmpMax1955.fit)
spplot(JanTmpMax1955krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





# AUGUST 1955 #



#Temperature values for Auguary 1955 in Texas
AugTmpMax1955 = readDataTemp("1955/Aug/1955-Aug-TMax.xlsx")

#data frame with data for one day
AugTmpMax1955Subset = computeTempAverage(AugTmpMax1955)
AugTmpMax1955Subset = coordinateTransformTemp(AugTmpMax1955Subset)

#statistics
AugTmpMax1955Stats = statisticsCompute(AugTmpMax1955Subset$meanTemp)

# variogram for the one day
vAugTmpMax1955 = variogram(meanTemp ~ LATITUDE+LONGITUDE, AugTmpMax1955Subset)
plot(vAugTmpMax1955)

#fitting variogram
vAugTmpMax1955.fit = fit.variogram(vAugTmpMax1955, vgm(2, 'Lin', 0, 1))
plot(vAugTmpMax1955 ,vAugTmpMax1955.fit)

#kriging
AugTmpMax1955krige1 = krige(meanTemp ~ 1, AugTmpMax1955Subset, texasGrid, vAugTmpMax1955.fit)
spplot(AugTmpMax1955krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))
```

```{r, echo=FALSE,  message=FALSE, warning=FALSE}
#functions for the kriging interpolation of the precipitation data
computePrcpAverage = function(dataFrame){
  dataFrameSubset = mutate(dataFrame, month = format(DATE, "%m"))
  dataFrameMean = group_by(dataFrameSubset, STATION, NAME, LATITUDE, LONGITUDE, ELEVATION, month) %>% summarise(meanPrcp = mean(PRCP, na.rm = T))
  dataFrameMean$meanPrcp = round(dataFrameMean$meanPrcp, digits=4)
  dataFrameMean = dataFrameMean[!rowSums(dataFrameMean[7]>15),]
  return(dataFrameMean)
}

coordinateTransformPrcp = function(dataSet){
  coordinates(dataSet) = dataSet[c(4,3)]
  proj4string(dataSet) = CRS("+init=epsg:4326")
  dataSet = spTransform(dataSet, CRS("+proj=utm +zone=13 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))
  return(dataSet)
}

readDataPrcp = function(path){
  rawData = read_xlsx(path)
  rawData$LATITUDE = as.numeric(rawData$LATITUDE)
  rawData$LONGITUDE = as.numeric(rawData$LONGITUDE)
  return(rawData)
}
```

```{r, echo=FALSE, fig.show="hide",  message=FALSE, warning=FALSE, results="hide"}
#analysis and kriging interpolation of the precipitation data
setwd("data")
                                # JANUARY 2015 #

# Precipitation values for January 2015 in Texas
JanPrcp2015 = readDataPrcp("2015/Jan/2015-Jan-Prcp.xlsx")

#data frame with data for one day
JanPrcp2015Subset = computePrcpAverage(JanPrcp2015)
JanPrcp2015Subset = coordinateTransformPrcp(JanPrcp2015Subset)

#statistics
JanPrcp2015Stats = statisticsCompute(JanPrcp2015Subset$meanPrcp)

# variogram for the one day
vJanPrcp2015 = variogram(meanPrcp ~ LATITUDE+LONGITUDE, JanPrcp2015Subset)
plot(vJanPrcp2015)

#fitting variogram
vJanPrcp2015.fit = fit.variogram(vJanPrcp2015, vgm('Exp'))
plot(vJanPrcp2015 ,vJanPrcp2015.fit)

#kriging
JanPrcp2015krige1 = krige(meanPrcp ~ 1, JanPrcp2015Subset, texasGrid, vJanPrcp2015.fit)
spplot(JanPrcp2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





  
                                  # AUGUST 2015 #

# Precipitation values for Auguary 2015 in Texas
AugPrcp2015 = readDataPrcp("2015/Aug/2015-Aug-Prcp.xlsx")

#data frame with data for one day
AugPrcp2015Subset = computePrcpAverage(AugPrcp2015)
AugPrcp2015Subset = coordinateTransformPrcp(AugPrcp2015Subset)

#statistics
AugPrcp2015Stats = statisticsCompute(AugPrcp2015Subset$meanPrcp)

# variogram for the one day
vAugPrcp2015 = variogram(meanPrcp ~ LATITUDE+LONGITUDE, AugPrcp2015Subset)
plot(vAugPrcp2015)

#fitting variogram
vAugPrcp2015.fit = fit.variogram(vAugPrcp2015, vgm('Exp'))
plot(vAugPrcp2015 ,vAugPrcp2015.fit)

#kriging
AugPrcp2015krige1 = krige(meanPrcp ~ 1, AugPrcp2015Subset, texasGrid, vAugPrcp2015.fit)
spplot(AugPrcp2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






                                      # JANUARY 2005 #

# Precipitation values for January 2005 in Texas
JanPrcp2005 = readDataPrcp("2005/Jan/2005-Jan-Prcp.xlsx")

#data frame with data for one day
JanPrcp2005Subset = computePrcpAverage(JanPrcp2005)
JanPrcp2005Subset = coordinateTransformPrcp(JanPrcp2005Subset)

#statistics
JanPrcp2005Stats = statisticsCompute(JanPrcp2005Subset$meanPrcp)

# variogram for the one day
vJanPrcp2005 = variogram(meanPrcp ~ LATITUDE+LONGITUDE, JanPrcp2005Subset)
plot(vJanPrcp2005)

#fitting variogram
vJanPrcp2005.fit = fit.variogram(vJanPrcp2005, vgm('Exp'))
plot(vJanPrcp2005 ,vJanPrcp2005.fit)

#kriging
JanPrcp2005krige1 = krige(meanPrcp ~ 1, JanPrcp2005Subset, texasGrid, vJanPrcp2005.fit)
spplot(JanPrcp2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






# AUGUST 2005 #

# Precipitation values for Auguary 2005 in Texas
AugPrcp2005 = readDataPrcp("2005/Aug/2005-Aug-Prcp.xlsx")

#data frame with data for one day
AugPrcp2005Subset = computePrcpAverage(AugPrcp2005)
AugPrcp2005Subset = coordinateTransformPrcp(AugPrcp2005Subset)

#statistics
AugPrcp2005Stats = statisticsCompute(AugPrcp2005Subset$meanPrcp)

# variogram for the one day
vAugPrcp2005 = variogram(meanPrcp ~ LATITUDE+LONGITUDE, AugPrcp2005Subset)
plot(vAugPrcp2005)

#fitting variogram
vAugPrcp2005.fit = fit.variogram(vAugPrcp2005, vgm('Exp'))
plot(vAugPrcp2005 ,vAugPrcp2005.fit)

#kriging
AugPrcp2005krige1 = krige(meanPrcp ~ 1, AugPrcp2005Subset, texasGrid, vAugPrcp2005.fit)
spplot(AugPrcp2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





                                        # JANUARY 1995 #

# Precipitation values for January 1995 in Texas
JanPrcp1995 = readDataPrcp("1995/Jan/1995-Jan-Prcp.xlsx")

#data frame with data for one day
JanPrcp1995Subset = computePrcpAverage(JanPrcp1995)
JanPrcp1995Subset = coordinateTransformPrcp(JanPrcp1995Subset)

#statistics
JanPrcp1995Stats = statisticsCompute(JanPrcp1995Subset$meanPrcp)

# variogram for the one day
vJanPrcp1995 = variogram(meanPrcp ~ LATITUDE+LONGITUDE, JanPrcp1995Subset)
plot(vJanPrcp1995)

#fitting variogram
vJanPrcp1995.fit = fit.variogram(vJanPrcp1995, vgm('Exp'))
plot(vJanPrcp1995 ,vJanPrcp1995.fit)

#kriging
JanPrcp1995krige1 = krige(meanPrcp ~ 1, JanPrcp1995Subset, texasGrid, vJanPrcp1995.fit)
spplot(JanPrcp1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






# AUGUST 1995 #

# Precipitation values for Auguary 1995 in Texas
AugPrcp1995 = readDataPrcp("1995/Aug/1995-Aug-Prcp.xlsx")

#data frame with data for one day
AugPrcp1995Subset = computePrcpAverage(AugPrcp1995)
AugPrcp1995Subset = coordinateTransformPrcp(AugPrcp1995Subset)

#statistics
AugPrcp1995Stats = statisticsCompute(AugPrcp1995Subset$meanPrcp)

# variogram for the one day
vAugPrcp1995 = variogram(meanPrcp ~ LATITUDE+LONGITUDE, AugPrcp1995Subset)
plot(vAugPrcp1995)

#fitting variogram
vAugPrcp1995.fit = fit.variogram(vAugPrcp1995, vgm('Exp'))
plot(vAugPrcp1995 ,vAugPrcp1995.fit)

#kriging
AugPrcp1995krige1 = krige(meanPrcp ~ 1, AugPrcp1995Subset, texasGrid, vAugPrcp1995.fit)
spplot(AugPrcp1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





                                          # JANUARY 1985 #

# Precipitation values for January 1985 in Texas
JanPrcp1985 = readDataPrcp("1985/Jan/1985-Jan-Prcp.xlsx")

#data frame with data for one day
JanPrcp1985Subset = computePrcpAverage(JanPrcp1985)
JanPrcp1985Subset = coordinateTransformPrcp(JanPrcp1985Subset)

#statistics
JanPrcp1985Stats = statisticsCompute(JanPrcp1985Subset$meanPrcp)

# variogram for the one day
vJanPrcp1985 = variogram(meanPrcp ~ LATITUDE+LONGITUDE, JanPrcp1985Subset)
plot(vJanPrcp1985)

#fitting variogram
vJanPrcp1985.fit = fit.variogram(vJanPrcp1985, vgm('Exp'))
plot(vJanPrcp1985 ,vJanPrcp1985.fit)

#kriging
JanPrcp1985krige1 = krige(meanPrcp ~ 1, JanPrcp1985Subset, texasGrid, vJanPrcp1985.fit)
spplot(JanPrcp1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






# AUGUST 1985 #

# Precipitation values for Auguary 1985 in Texas
AugPrcp1985 = readDataPrcp("1985/Aug/1985-Aug-Prcp.xlsx")

#data frame with data for one day
AugPrcp1985Subset = computePrcpAverage(AugPrcp1985)
AugPrcp1985Subset = coordinateTransformPrcp(AugPrcp1985Subset)

#statistics
AugPrcp1985Stats = statisticsCompute(AugPrcp1985Subset$meanPrcp)

# variogram for the one day
vAugPrcp1985 = variogram(meanPrcp ~ LATITUDE+LONGITUDE, AugPrcp1985Subset)
plot(vAugPrcp1985)

#fitting variogram
vAugPrcp1985.fit = fit.variogram(vAugPrcp1985, vgm('Exp'))
plot(vAugPrcp1985 ,vAugPrcp1985.fit)

#kriging
AugPrcp1985krige1 = krige(meanPrcp ~ 1, AugPrcp1985Subset, texasGrid, vAugPrcp1985.fit)
spplot(AugPrcp1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





                                                  # JANUARY 1975 #

# Precipitation values for January 1975 in Texas
JanPrcp1975 = readDataPrcp("1975/Jan/1975-Jan-Prcp.xlsx")

#data frame with data for one day
JanPrcp1975Subset = computePrcpAverage(JanPrcp1975)
JanPrcp1975Subset = coordinateTransformPrcp(JanPrcp1975Subset)

#statistics
JanPrcp1975Stats = statisticsCompute(JanPrcp1975Subset$meanPrcp)

# variogram for the one day
vJanPrcp1975 = variogram(meanPrcp ~ LATITUDE+LONGITUDE, JanPrcp1975Subset)
plot(vJanPrcp1975)

#fitting variogram
vJanPrcp1975.fit = fit.variogram(vJanPrcp1975, vgm('Exp'))
plot(vJanPrcp1975 ,vJanPrcp1975.fit)

#kriging
JanPrcp1975krige1 = krige(meanPrcp ~ 1, JanPrcp1975Subset, texasGrid, vJanPrcp1975.fit)
spplot(JanPrcp1975krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






# AUGUST 1975 #

# Precipitation values for Auguary 1975 in Texas
AugPrcp1975 = readDataPrcp("1975/Aug/1975-Aug-Prcp.xlsx")

#data frame with data for one day
AugPrcp1975Subset = computePrcpAverage(AugPrcp1975)
AugPrcp1975Subset = coordinateTransformPrcp(AugPrcp1975Subset)

#statistics
AugPrcp1975Stats = statisticsCompute(AugPrcp1975Subset$meanPrcp)

# variogram for the one day
vAugPrcp1975 = variogram(meanPrcp ~ LATITUDE+LONGITUDE, AugPrcp1975Subset)
plot(vAugPrcp1975)

#fitting variogram
vAugPrcp1975.fit = fit.variogram(vAugPrcp1975, vgm('Exp'))
plot(vAugPrcp1975 ,vAugPrcp1975.fit)

#kriging
AugPrcp1975krige1 = krige(meanPrcp ~ 1, AugPrcp1975Subset, texasGrid, vAugPrcp1975.fit)
spplot(AugPrcp1975krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





                                    # JANUARY 1965 #

# Precipitation values for January 1965 in Texas
JanPrcp1965 = readDataPrcp("1965/Jan/1965-Jan-Prcp.xlsx")

#data frame with data for one day
JanPrcp1965Subset = computePrcpAverage(JanPrcp1965)
JanPrcp1965Subset = coordinateTransformPrcp(JanPrcp1965Subset)

#statistics
JanPrcp1965Stats = statisticsCompute(JanPrcp1965Subset$meanPrcp)

# variogram for the one day
vJanPrcp1965 = variogram(meanPrcp ~ LATITUDE+LONGITUDE, JanPrcp1965Subset)
plot(vJanPrcp1965)

#fitting variogram
vJanPrcp1965.fit = fit.variogram(vJanPrcp1965, vgm('Exp'))
plot(vJanPrcp1965 ,vJanPrcp1965.fit)

#kriging
JanPrcp1965krige1 = krige(meanPrcp ~ 1, JanPrcp1965Subset, texasGrid, vJanPrcp1965.fit)
spplot(JanPrcp1965krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






# AUGUST 1965 #

# Precipitation values for Auguary 1965 in Texas
AugPrcp1965 = readDataPrcp("1965/Aug/1965-Aug-Prcp.xlsx")

#data frame with data for one day
AugPrcp1965Subset = computePrcpAverage(AugPrcp1965)
AugPrcp1965Subset = coordinateTransformPrcp(AugPrcp1965Subset)

#statistics
AugPrcp1965Stats = statisticsCompute(AugPrcp1965Subset$meanPrcp)

# variogram for the one day
vAugPrcp1965 = variogram(meanPrcp ~ LATITUDE+LONGITUDE, AugPrcp1965Subset)
plot(vAugPrcp1965)

#fitting variogram
vAugPrcp1965.fit = fit.variogram(vAugPrcp1965, vgm('Exp'))
plot(vAugPrcp1965 ,vAugPrcp1965.fit)

#kriging
AugPrcp1965krige1 = krige(meanPrcp ~ 1, AugPrcp1965Subset, texasGrid, vAugPrcp1965.fit)
spplot(AugPrcp1965krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





                                    # JANUARY 1955 #

# Precipitation values for January 1955 in Texas
JanPrcp1955 = readDataPrcp("1955/Jan/1955-Jan-Prcp.xlsx")

#data frame with data for one day
JanPrcp1955Subset = computePrcpAverage(JanPrcp1955)
JanPrcp1955Subset = coordinateTransformPrcp(JanPrcp1955Subset)

#statistics
JanPrcp1955Stats = statisticsCompute(JanPrcp1955Subset$meanPrcp)

# variogram for the one day
vJanPrcp1955 = variogram(meanPrcp ~ LATITUDE+LONGITUDE, JanPrcp1955Subset)
plot(vJanPrcp1955)

#fitting variogram
vJanPrcp1955.fit = fit.variogram(vJanPrcp1955, vgm('Exp'))
plot(vJanPrcp1955 ,vJanPrcp1955.fit)

#kriging
JanPrcp1955krige1 = krige(meanPrcp ~ 1, JanPrcp1955Subset, texasGrid, vJanPrcp1955.fit)
spplot(JanPrcp1955krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






# AUGUST 1955 #

# Precipitation values for Auguary 1955 in Texas
AugPrcp1955 = readDataPrcp("1955/Aug/1955-Aug-Prcp.xlsx")

#data frame with data for one day
AugPrcp1955Subset = computePrcpAverage(AugPrcp1955)
AugPrcp1955Subset = coordinateTransformPrcp(AugPrcp1955Subset)

#statistics
AugPrcp1955Stats = statisticsCompute(AugPrcp1955Subset$meanPrcp)

# variogram for the one day
vAugPrcp1955 = variogram(meanPrcp ~ LATITUDE+LONGITUDE, AugPrcp1955Subset)
plot(vAugPrcp1955)

#fitting variogram
vAugPrcp1955.fit = fit.variogram(vAugPrcp1955, vgm('Exp'))
plot(vAugPrcp1955 ,vAugPrcp1955.fit)

#kriging
AugPrcp1955krige1 = krige(meanPrcp ~ 1, AugPrcp1955Subset, texasGrid, vAugPrcp1955.fit)
spplot(AugPrcp1955krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))
```

```{r, echo=FALSE,  message=FALSE, warning=FALSE}
#functions for the kriging interpolation of the evaporation data
computeEvapAverage = function(dataFrame){
  dataFrameSubset = mutate(dataFrame, month = format(DATE, "%m"))
  dataFrameMean = group_by(dataFrameSubset, STATION, NAME, LATITUDE, LONGITUDE, ELEVATION, month) %>% summarise(meanEvap = mean(EVAP, na.rm = T))
  dataFrameMean$meanEvap = round(dataFrameMean$meanEvap, digits=4)
  dataFrameMean = dataFrameMean[!rowSums(dataFrameMean[7]>30),]
  return(dataFrameMean)
}

coordinateTransformEvap = function(dataSet){
  coordinates(dataSet) = dataSet[c(4,3)]
  proj4string(dataSet) = CRS("+init=epsg:4326")
  dataSet = spTransform(dataSet, CRS("+proj=utm +zone=13 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))
  return(dataSet)
}

readDataEvap = function(path){
  rawData = read_xlsx(path)
  rawData$LATITUDE = as.numeric(rawData$LATITUDE)
  rawData$LONGITUDE = as.numeric(rawData$LONGITUDE)
  return(rawData)
}
```

```{r, echo=FALSE, fig.show="hide",  message=FALSE, warning=FALSE, results="hide"}
#analysis and kriging interpolation of the evaporation data
setwd("data")
                                      # JANUARY 2015 #

# Evaporation values for January 2015 in Texas
JanEvap2015 = readDataEvap("2015/Jan/2015-Jan-Evap.xlsx")

#data frame with data for one day
JanEvap2015Subset = computeEvapAverage(JanEvap2015)
JanEvap2015Subset = coordinateTransformEvap(JanEvap2015Subset)

#statistics
JanEvap2015Stats = statisticsCompute(JanEvap2015Subset$meanEvap)

# variogram for the one day
vJanEvap2015 = variogram(meanEvap ~ LATITUDE+LONGITUDE, JanEvap2015Subset)
plot(vJanEvap2015)

#fitting variogram
vJanEvap2015.fit = fit.variogram(vJanEvap2015, vgm(1.5, 'Lin', 0))
plot(vJanEvap2015 ,vJanEvap2015.fit)

#kriging
JanEvap2015krige1 = krige(meanEvap ~ 1, JanEvap2015Subset, texasGrid, vJanEvap2015.fit)
spplot(JanEvap2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






                                            # AUGUST 2015 #

# Evaporation values for Auguary 2015 in Texas
AugEvap2015 = readDataEvap("2015/Aug/2015-Aug-Evap.xlsx")

#data frame with data for one day
AugEvap2015Subset = computeEvapAverage(AugEvap2015)
AugEvap2015Subset = coordinateTransformEvap(AugEvap2015Subset)

#statistics
AugEvap2015Stats = statisticsCompute(AugEvap2015Subset$meanEvap)

# variogram for the one day
vAugEvap2015 = variogram(meanEvap ~ LATITUDE+LONGITUDE, AugEvap2015Subset)
plot(vAugEvap2015)

#fitting variogram
vAugEvap2015.fit = fit.variogram(vAugEvap2015, vgm(1.5, 'Lin', 0))
plot(vAugEvap2015 ,vAugEvap2015.fit)

#kriging
AugEvap2015krige1 = krige(meanEvap ~ 1, AugEvap2015Subset, texasGrid, vAugEvap2015.fit)
spplot(AugEvap2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))







                                          # JANUARY 2005 #

# Evaporation values for January 2005 in Texas
JanEvap2005 = readDataEvap("2005/Jan/2005-Jan-Evap.xlsx")

#data frame with data for one day
JanEvap2005Subset = computeEvapAverage(JanEvap2005)
JanEvap2005Subset = coordinateTransformEvap(JanEvap2005Subset)

#statistics
JanEvap2005Stats = statisticsCompute(JanEvap2005Subset$meanEvap)

# variogram for the one day
vJanEvap2005 = variogram(meanEvap ~ LATITUDE+LONGITUDE, JanEvap2005Subset)
plot(vJanEvap2005)

#fitting variogram
vJanEvap2005.fit = fit.variogram(vJanEvap2005, vgm(1.5, 'Lin', 0))
plot(vJanEvap2005 ,vJanEvap2005.fit)

#kriging
JanEvap2005krige1 = krige(meanEvap ~ 1, JanEvap2005Subset, texasGrid, vJanEvap2005.fit)
spplot(JanEvap2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






# AUGUST 2005 #

# Evaporation values for Auguary 2005 in Texas
AugEvap2005 = readDataEvap("2005/Aug/2005-Aug-Evap.xlsx")

#data frame with data for one day
AugEvap2005Subset = computeEvapAverage(AugEvap2005)
AugEvap2005Subset = coordinateTransformEvap(AugEvap2005Subset)

#statistics
AugEvap2005Stats = statisticsCompute(AugEvap2005Subset$meanEvap)

# variogram for the one day
vAugEvap2005 = variogram(meanEvap ~ LATITUDE+LONGITUDE, AugEvap2005Subset)
plot(vAugEvap2005)

#fitting variogram
vAugEvap2005.fit = fit.variogram(vAugEvap2005, vgm(1.5, 'Lin', 0))
plot(vAugEvap2005 ,vAugEvap2005.fit)

#kriging
AugEvap2005krige1 = krige(meanEvap ~ 1, AugEvap2005Subset, texasGrid, vAugEvap2005.fit)
spplot(AugEvap2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






                                            # JANUARY 1995 #

# Evaporation values for January 1995 in Texas
JanEvap1995 = readDataEvap("1995/Jan/1995-Jan-Evap.xlsx")

#data frame with data for one day
JanEvap1995Subset = computeEvapAverage(JanEvap1995)
JanEvap1995Subset = coordinateTransformEvap(JanEvap1995Subset)

#statistics
JanEvap1995Stats = statisticsCompute(JanEvap1995Subset$meanEvap)

# variogram for the one day
vJanEvap1995 = variogram(meanEvap ~ LATITUDE+LONGITUDE, JanEvap1995Subset)
plot(vJanEvap1995)

#fitting variogram
vJanEvap1995.fit = fit.variogram(vJanEvap1995, vgm(1.5, 'Lin', 0))
plot(vJanEvap1995 ,vJanEvap1995.fit)

#kriging
JanEvap1995krige1 = krige(meanEvap ~ 1, JanEvap1995Subset, texasGrid, vJanEvap1995.fit)
spplot(JanEvap1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






# AUGUST 1995 #

# Evaporation values for Auguary 1995 in Texas
AugEvap1995 = readDataEvap("1995/Aug/1995-Aug-Evap.xlsx")

#data frame with data for one day
AugEvap1995Subset = computeEvapAverage(AugEvap1995)
AugEvap1995Subset = coordinateTransformEvap(AugEvap1995Subset)

#statistics
AugEvap1995Stats = statisticsCompute(AugEvap1995Subset$meanEvap)

# variogram for the one day
vAugEvap1995 = variogram(meanEvap ~ LATITUDE+LONGITUDE, AugEvap1995Subset)
plot(vAugEvap1995)

#fitting variogram
vAugEvap1995.fit = fit.variogram(vAugEvap1995, vgm(1.5, 'Lin', 0))
plot(vAugEvap1995 ,vAugEvap1995.fit)

#kriging
AugEvap1995krige1 = krige(meanEvap ~ 1, AugEvap1995Subset, texasGrid, vAugEvap1995.fit)
spplot(AugEvap1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))








                                          # JANUARY 1985 #

# Evaporation values for January 1985 in Texas
JanEvap1985 = readDataEvap("1985/Jan/1985-Jan-Evap.xlsx")

#data frame with data for one day
JanEvap1985Subset = computeEvapAverage(JanEvap1985)
JanEvap1985Subset = coordinateTransformEvap(JanEvap1985Subset)

#statistics
JanEvap1985Stats = statisticsCompute(JanEvap1985Subset$meanEvap)

# variogram for the one day
vJanEvap1985 = variogram(meanEvap ~ LATITUDE+LONGITUDE, JanEvap1985Subset)
plot(vJanEvap1985)

#fitting variogram
vJanEvap1985.fit = fit.variogram(vJanEvap1985, vgm(1.5, 'Lin', 0))
plot(vJanEvap1985 ,vJanEvap1985.fit)

#kriging
JanEvap1985krige1 = krige(meanEvap ~ 1, JanEvap1985Subset, texasGrid, vJanEvap1985.fit)
spplot(JanEvap1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






# AUGUST 1985 #

# Evaporation values for Auguary 1985 in Texas
AugEvap1985 = readDataEvap("1985/Aug/1985-Aug-Evap.xlsx")

#data frame with data for one day
AugEvap1985Subset = computeEvapAverage(AugEvap1985)
AugEvap1985Subset = coordinateTransformEvap(AugEvap1985Subset)

#statistics
AugEvap1985Stats = statisticsCompute(AugEvap1985Subset$meanEvap)

# variogram for the one day
vAugEvap1985 = variogram(meanEvap ~ LATITUDE+LONGITUDE, AugEvap1985Subset)
plot(vAugEvap1985)

#fitting variogram
vAugEvap1985.fit = fit.variogram(vAugEvap1985, vgm(1.5, 'Lin', 0))
plot(vAugEvap1985 ,vAugEvap1985.fit)

#kriging
AugEvap1985krige1 = krige(meanEvap ~ 1, AugEvap1985Subset, texasGrid, vAugEvap1985.fit)
spplot(AugEvap1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))







                                          # JANUARY 1975 #

# Evaporation values for January 1975 in Texas
JanEvap1975 = readDataEvap("1975/Jan/1975-Jan-Evap.xlsx")

#data frame with data for one day
JanEvap1975Subset = computeEvapAverage(JanEvap1975)
JanEvap1975Subset = coordinateTransformEvap(JanEvap1975Subset)

#statistics
JanEvap1975Stats = statisticsCompute(JanEvap1975Subset$meanEvap)

# variogram for the one day
vJanEvap1975 = variogram(meanEvap ~ LATITUDE+LONGITUDE, JanEvap1975Subset)
plot(vJanEvap1975)

#fitting variogram
vJanEvap1975.fit = fit.variogram(vJanEvap1975, vgm(1.5, 'Lin', 0))
plot(vJanEvap1975 ,vJanEvap1975.fit)

#kriging
JanEvap1975krige1 = krige(meanEvap ~ 1, JanEvap1975Subset, texasGrid, vJanEvap1975.fit)
spplot(JanEvap1975krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






# AUGUST 1975 #

# Evaporation values for Auguary 1975 in Texas
AugEvap1975 = readDataEvap("1975/Aug/1975-Aug-Evap.xlsx")

#data frame with data for one day
AugEvap1975Subset = computeEvapAverage(AugEvap1975)
AugEvap1975Subset = coordinateTransformEvap(AugEvap1975Subset)

#statistics
AugEvap1975Stats = statisticsCompute(AugEvap1975Subset$meanEvap)

# variogram for the one day
vAugEvap1975 = variogram(meanEvap ~ LATITUDE+LONGITUDE, AugEvap1975Subset)
plot(vAugEvap1975)

#fitting variogram
vAugEvap1975.fit = fit.variogram(vAugEvap1975, vgm(1.5, 'Lin', 0))
plot(vAugEvap1975 ,vAugEvap1975.fit)

#kriging
AugEvap1975krige1 = krige(meanEvap ~ 1, AugEvap1975Subset, texasGrid, vAugEvap1975.fit)
spplot(AugEvap1975krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






                                          # JANUARY 1965 #

# Evaporation values for January 1965 in Texas
JanEvap1965 = readDataEvap("1965/Jan/1965-Jan-Evap.xlsx")

#data frame with data for one day
JanEvap1965Subset = computeEvapAverage(JanEvap1965)
JanEvap1965Subset = coordinateTransformEvap(JanEvap1965Subset)

#statistics
JanEvap1965Stats = statisticsCompute(JanEvap1965Subset$meanEvap)

# variogram for the one day
vJanEvap1965 = variogram(meanEvap ~ LATITUDE+LONGITUDE, JanEvap1965Subset)
plot(vJanEvap1965)

#fitting variogram
vJanEvap1965.fit = fit.variogram(vJanEvap1965, vgm(1.5, 'Lin', 0))
plot(vJanEvap1965 ,vJanEvap1965.fit)

#kriging
JanEvap1965krige1 = krige(meanEvap ~ 1, JanEvap1965Subset, texasGrid, vJanEvap1965.fit)
spplot(JanEvap1965krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






# AUGUST 1965 #

# Evaporation values for Auguary 1965 in Texas
AugEvap1965 = readDataEvap("1965/Aug/1965-Aug-Evap.xlsx")

#data frame with data for one day
AugEvap1965Subset = computeEvapAverage(AugEvap1965)
AugEvap1965Subset = coordinateTransformEvap(AugEvap1965Subset)

#statistics
AugEvap1965Stats = statisticsCompute(AugEvap1965Subset$meanEvap)

# variogram for the one day
vAugEvap1965 = variogram(meanEvap ~ LATITUDE+LONGITUDE, AugEvap1965Subset)
plot(vAugEvap1965)

#fitting variogram
vAugEvap1965.fit = fit.variogram(vAugEvap1965, vgm(1.5, 'Lin', 0))
plot(vAugEvap1965 ,vAugEvap1965.fit)

#kriging
AugEvap1965krige1 = krige(meanEvap ~ 1, AugEvap1965Subset, texasGrid, vAugEvap1965.fit)
spplot(AugEvap1965krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))







                                          # JANUARY 1955 #

# Evaporation values for January 1955 in Texas
JanEvap1955 = readDataEvap("1955/Jan/1955-Jan-Evap.xlsx")

#data frame with data for one day
JanEvap1955Subset = computeEvapAverage(JanEvap1955)
JanEvap1955Subset = coordinateTransformEvap(JanEvap1955Subset)

#statistics
JanEvap1955Stats = statisticsCompute(JanEvap1955Subset$meanEvap)

# variogram for the one day
vJanEvap1955 = variogram(meanEvap ~ LATITUDE+LONGITUDE, JanEvap1955Subset)
plot(vJanEvap1955)

#fitting variogram
vJanEvap1955.fit = fit.variogram(vJanEvap1955, vgm(1.5, 'Lin', 0))
plot(vJanEvap1955 ,vJanEvap1955.fit)

#kriging
JanEvap1955krige1 = krige(meanEvap ~ 1, JanEvap1955Subset, texasGrid, vJanEvap1955.fit)
spplot(JanEvap1955krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






# AUGUST 1955 #

# Evaporation values for Auguary 1955 in Texas
AugEvap1955 = readDataEvap("1955/Aug/1955-Aug-Evap.xlsx")

#data frame with data for one day
AugEvap1955Subset = computeEvapAverage(AugEvap1955)
AugEvap1955Subset = coordinateTransformEvap(AugEvap1955Subset)

#statistics
AugEvap1955Stats = statisticsCompute(AugEvap1955Subset$meanEvap)

# variogram for the one day
vAugEvap1955 = variogram(meanEvap ~ LATITUDE+LONGITUDE, AugEvap1955Subset)
plot(vAugEvap1955)

#fitting variogram
vAugEvap1955.fit = fit.variogram(vAugEvap1955, vgm(1.5, 'Lin', 0))
plot(vAugEvap1955 ,vAugEvap1955.fit)

#kriging
AugEvap1955krige1 = krige(meanEvap ~ 1, AugEvap1955Subset, texasGrid, vAugEvap1955.fit)
spplot(AugEvap1955krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))
```

```{r, echo=FALSE,  message=FALSE, warning=FALSE}
#functions for the kriging interpolation of the wind data
computeWindAverage = function(dataFrame){
  dataFrameSubset = mutate(dataFrame, month = format(DATE, "%m"))
  dataFrameMean = group_by(dataFrameSubset, STATION, NAME, LATITUDE, LONGITUDE, ELEVATION, month) %>% summarise(meanWind = mean(AWND, na.rm = T))
  dataFrameMean$meanWind = round(dataFrameMean$meanWind, digits=4)
  return(dataFrameMean)
}

coordinateTransformWind = function(dataSet){
  coordinates(dataSet) = dataSet[c(4,3)]
  proj4string(dataSet) = CRS("+init=epsg:4326")
  dataSet = spTransform(dataSet, CRS("+proj=utm +zone=13 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))
  return(dataSet)
}

readDataWind = function(path){
  rawData = read_xlsx(path)
  rawData$LATITUDE = as.numeric(rawData$LATITUDE)
  rawData$LONGITUDE = as.numeric(rawData$LONGITUDE)
  return(rawData)
}
```

```{r, echo=FALSE, fig.show="hide",  message=FALSE, warning=FALSE, results="hide"}
#analysis and kriging interpolation of the wind data
setwd("data")
                                      # JANUARY 2015 #

# Average wind speed values for January 2015 in Texas
JanWind2015 = readDataWind("2015/Jan/2015-Jan-Wind.xlsx")

#data frame with data for one day
JanWind2015Subset = computeWindAverage(JanWind2015)
JanWind2015Subset = coordinateTransformWind(JanWind2015Subset)

#statistics
JanWind2015Stats = statisticsCompute(JanWind2015Subset$meanWind)

# variogram for the one day
vJanWind2015 = variogram(meanWind ~ LATITUDE+LONGITUDE, JanWind2015Subset)
plot(vJanWind2015)

#fitting variogram
vJanWind2015.fit = fit.variogram(vJanWind2015, vgm(1, 'Lin', 0))
plot(vJanWind2015 ,vJanWind2015.fit)

#kriging
JanWind2015krige1 = krige(meanWind ~ 1, JanWind2015Subset, texasGrid, vJanWind2015.fit)
spplot(JanWind2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))







                                            # AUGUST 2015 #

# Average wind speed values for Auguary 2015 in Texas
AugWind2015 = readDataWind("2015/Aug/2015-Aug-Wind.xlsx")

#data frame with data for one day
AugWind2015Subset = computeWindAverage(AugWind2015)
AugWind2015Subset = coordinateTransformWind(AugWind2015Subset)

#statistics
AugWind2015Stats = statisticsCompute(AugWind2015Subset$meanWind)

# variogram for the one day
vAugWind2015 = variogram(meanWind ~ LATITUDE+LONGITUDE, AugWind2015Subset)
plot(vAugWind2015)

#fitting variogram
vAugWind2015.fit = fit.variogram(vAugWind2015, vgm(1, 'Lin', 0))
plot(vAugWind2015 ,vAugWind2015.fit)

#kriging
AugWind2015krige1 = krige(meanWind ~ 1, AugWind2015Subset, texasGrid, vAugWind2015.fit)
spplot(AugWind2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))





                                              # JANUARY 2005 #

# Average wind speed values for January 2005 in Texas
JanWind2005 = readDataWind("2005/Jan/2005-Jan-Wind.xlsx")

#data frame with data for one day
JanWind2005Subset = computeWindAverage(JanWind2005)
JanWind2005Subset = coordinateTransformWind(JanWind2005Subset)

#statistics
JanWind2005Stats = statisticsCompute(JanWind2005Subset$meanWind)

# variogram for the one day
vJanWind2005 = variogram(meanWind ~ LATITUDE+LONGITUDE, JanWind2005Subset)
plot(vJanWind2005)

#fitting variogram
vJanWind2005.fit = fit.variogram(vJanWind2005, vgm(1, 'Lin', 0))
plot(vJanWind2005 ,vJanWind2005.fit)

#kriging
JanWind2005krige1 = krige(meanWind ~ 1, JanWind2005Subset, texasGrid, vJanWind2005.fit)
spplot(JanWind2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))







# AUGUST 2005 #

# Average wind speed values for Auguary 2005 in Texas
AugWind2005 = readDataWind("2005/Aug/2005-Aug-Wind.xlsx")

#data frame with data for one day
AugWind2005Subset = computeWindAverage(AugWind2005)
AugWind2005Subset = coordinateTransformWind(AugWind2005Subset)

#statistics
AugWind2005Stats = statisticsCompute(AugWind2005Subset$meanWind)

# variogram for the one day
vAugWind2005 = variogram(meanWind ~ LATITUDE+LONGITUDE, AugWind2005Subset)
plot(vAugWind2005)

#fitting variogram
vAugWind2005.fit = fit.variogram(vAugWind2005, vgm(1, 'Lin', 0))
plot(vAugWind2005 ,vAugWind2005.fit)

#kriging
AugWind2005krige1 = krige(meanWind ~ 1, AugWind2005Subset, texasGrid, vAugWind2005.fit)
spplot(AugWind2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))








                                                # JANUARY 1995 #

# Average wind speed values for January 1995 in Texas
JanWind1995 = readDataWind("1995/Jan/1995-Jan-Wind.xlsx")

#data frame with data for one day
JanWind1995Subset = computeWindAverage(JanWind1995)
JanWind1995Subset = coordinateTransformWind(JanWind1995Subset)

#statistics
JanWind1995Stats = statisticsCompute(JanWind1995Subset$meanWind)

# variogram for the one day
vJanWind1995 = variogram(meanWind ~ LATITUDE+LONGITUDE, JanWind1995Subset)
plot(vJanWind1995)

#fitting variogram
vJanWind1995.fit = fit.variogram(vJanWind1995, vgm(1, 'Lin', 0))
plot(vJanWind1995 ,vJanWind1995.fit)

#kriging
JanWind1995krige1 = krige(meanWind ~ 1, JanWind1995Subset, texasGrid, vJanWind1995.fit)
spplot(JanWind1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))







# AUGUST 1995 #

# Average wind speed values for Auguary 1995 in Texas
AugWind1995 = readDataWind("1995/Aug/1995-Aug-Wind.xlsx")

#data frame with data for one day
AugWind1995Subset = computeWindAverage(AugWind1995)
AugWind1995Subset = coordinateTransformWind(AugWind1995Subset)

#statistics
AugWind1995Stats = statisticsCompute(AugWind1995Subset$meanWind)

# variogram for the one day
vAugWind1995 = variogram(meanWind ~ LATITUDE+LONGITUDE, AugWind1995Subset)
plot(vAugWind1995)

#fitting variogram
vAugWind1995.fit = fit.variogram(vAugWind1995, vgm(1, 'Lin', 0))
plot(vAugWind1995 ,vAugWind1995.fit)

#kriging
AugWind1995krige1 = krige(meanWind ~ 1, AugWind1995Subset, texasGrid, vAugWind1995.fit)
spplot(AugWind1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))






                                      # JANUARY 1985 #

# Average wind speed values for January 1985 in Texas
JanWind1985 = readDataWind("1985/Jan/1985-Jan-Wind.xlsx")

#data frame with data for one day
JanWind1985Subset = computeWindAverage(JanWind1985)
JanWind1985Subset = coordinateTransformWind(JanWind1985Subset)

#statistics
JanWind1985Stats = statisticsCompute(JanWind1985Subset$meanWind)

# variogram for the one day
vJanWind1985 = variogram(meanWind ~ LATITUDE+LONGITUDE, JanWind1985Subset)
plot(vJanWind1985)

#fitting variogram
vJanWind1985.fit = fit.variogram(vJanWind1985, vgm(1, 'Lin', 0))
plot(vJanWind1985 ,vJanWind1985.fit)

#kriging
JanWind1985krige1 = krige(meanWind ~ 1, JanWind1985Subset, texasGrid, vJanWind1985.fit)
spplot(JanWind1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))







# AUGUST 1985 #

# Average wind speed values for Auguary 1985 in Texas
AugWind1985 = readDataWind("1985/Aug/1985-Aug-Wind.xlsx")

#data frame with data for one day
AugWind1985Subset = computeWindAverage(AugWind1985)
AugWind1985Subset = coordinateTransformWind(AugWind1985Subset)

#statistics
AugWind1985Stats = statisticsCompute(AugWind1985Subset$meanWind)

# variogram for the one day
vAugWind1985 = variogram(meanWind ~ LATITUDE+LONGITUDE, AugWind1985Subset)
plot(vAugWind1985)

#fitting variogram
vAugWind1985.fit = fit.variogram(vAugWind1985, vgm(1, 'Lin', 0))
plot(vAugWind1985 ,vAugWind1985.fit)

#kriging
AugWind1985krige1 = krige(meanWind ~ 1, AugWind1985Subset, texasGrid, vAugWind1985.fit)
spplot(AugWind1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"))
```

## Abstract

We know about the climate change and its consequences since many years and a lot of research has been done in this field. The climate is a complex system compound of different elements like temperature, precipitation or wind. This paper aims to analyse the power of the climate change regarding these elements on the example of Texas. The effects of the climate change on air temperature, precipitation, wind speed and evaporation will be investigated. By analysing time series and kriging interpolations of data from weather stations across the state it showed out, that there is a remarkable climate change. The temperature is rising, precipitation slightly increases, wind speed slowed, and evaporation becomes weaker. This paper gives another evidence to the climate change and together with other examples we should take future political decisions with respect to research in this field.

## Introduction

Since Charles David Keeling presented his findings of an increase in the concentration of carbon dioxide in Earth's atmosphere, people started believing in a human influenced climate change (Keeling, 1960). Before that, the natural variability of the Earth's climate with changes into both extremes, cold and warm, over the last 800.000 years was known (Jouzel et al. 2007). Based on Keelings data and other climatic investigations it became clear that the current global warming is strongly human-induced. Since then, many started investigating the impact of the climate change on natural and human environment. Above all the IPCC, the Intergovernmental Panel on Climate Change, provides information about the human-induced climate change, the risks and consequences of it and the possibilities of a climate mitigation (IPCC, 2018). The summarized findings of their researcher groups show that there is a remarkable rise of temperature and a lot of significant changes in physical and biological systems in all parts of the earth (IPCC Working Group 1, 2007). Another working group shows the influence of the humans in the climate change (IPCC Working Group 2, 2007). 
To show how the climate change influences some of the most important climate elements: air temperature, precipitation, wind speed and evaporation, this paper investigates data gathered from weather stations all over Texas. Texas was chosen because it is a very large state (696,241 km2), is climatically interesting (containing three climate zones) and has a lot of easy accessible data (Texas Almanac, 2018). On top of that, Texas emits the most greenhouse gases of all states in the United States which makes it to a massive climate sinner (DallasNews, 2007). In addition to the book "The impact of global warming on Texas", by Schmandt et al., this paper takes a deeper look into data from single stations. In their work they gave an overview for the whole state regarding different climatic, biological and political factors and were able to show a remarkable climate change for precipitation and temperature (Schmandt et al., 2011). In this work the focus lies on single stations and interpolated data out of that. Additionally, it will be checked if there are differences between the climate change at the coast and at the interior and a forecast for the next years will be given.
This work starts by introducing the used data and how it was gathered. After that, the methods will be described which were used to analyse the data. This is followed by a presentation of the results and a discussion of the findings. It will be closed by a conclusion with an outlook to further research.


## Data

Two different types of data are used for the analysis:

* Daily time series data of four weather stations in Texas from 1950 until today
* Averaged data of all weather stations in Texas of January and August every ten years from 1955 on

The datasets involve the location of the station, the name and identifier of the station, the date of the observation and the measured value. The four measured climate elements are:

* Maximum air temperature
* Total precipitation
* Average wind speed
* Total evaporation from an evaporation pan

The four stations with the time series data were chosen to represent different parts of Texas. To test if there are differences between the coastal regions and the interior, one station for each of the four cardinal directions north, east, south and west were chosen.

```{r, echo=FALSE,  message=FALSE, warning=FALSE}
#plot of the locations of the four stations
angletonPlot = angletonTemp[1,c(3,4,5)]
coordinates(angletonPlot) = angletonPlot[c(3,2)]
BenbrookPlot = BenbrookTemp[1,c(3,4,5)]
coordinates(BenbrookPlot) = BenbrookPlot[c(3,2)]
fortStocktonPlot = fortStocktonTemp[1,c(3,4,5)]
coordinates(fortStocktonPlot) = fortStocktonPlot[c(3,2)]
MccookPlot = MccookTemp[1,c(3,4,5)]
coordinates(MccookPlot) = MccookPlot[c(3,2)]
map('state', 'texas')
title(main="Locations of the four used stations in Texas for the time series analysis")
plot(angletonPlot, add=TRUE, pch=15, col="blue")
text(angletonPlot, label=angletonPlot$NAME, pos=3, cex=.6)
plot(BenbrookPlot, add=TRUE, pch=15, col="blue")
text(BenbrookPlot, label=BenbrookPlot$NAME, pos=3, cex=.6)
plot(fortStocktonPlot, add=TRUE, pch=15, col="blue")
text(fortStocktonPlot, label=fortStocktonPlot$NAME, pos=3, cex=.6)
plot(MccookPlot, add=TRUE, pch=15, col="blue")
text(MccookPlot, label=MccookPlot$NAME, pos=3, cex=.6)
```

Since by the chosen stations only measured the total wind movement instead of average wind speed, this is used for the analysis. The total wind movement are the summed wind speeds for one day. The temporal amount of the time series reaches from 1950 until today. Not every series of data from the four stations covers the whole extent, but the chosen ones are the best fitting stations in sense of location and data coverage for this analysis. Missing data for single days is interpolated or left out if longer periods are missing.

```{r, echo=FALSE,  message=FALSE, warning=FALSE}
#example time series for every data type of the station in Mccook
plot(tmpMaxMccook, main="Daily Maximum Temperature in Mccook, TX")
plot(PrcpMccook, main="Daily Precipitation in Mccook, TX")
plot(EvapMccook, main="Daily Evaporation from an Evaporation Pan in Mccook, TX")
plot(WindMccook, main="Daily Total Wind Movement in Mccook, TX")
```

The data of entire Texas includes more than 4500 weather stations but not every station measured all presented climate elements. The values from January and August are averaged daily measurements to get more representative data. Missing data for single days will be omit in the average computation.

```{r}
#head of the averaged temperature dataset of Texas in January 2015 with UTM coordinates
head(JanTmpMax2015Subset[c(1,2,3,4,7)])

#head of the averaged precipitation dataset of Texas in January 2015 with UTM coordinates
head(JanPrcp2015Subset[c(1,2,3,4,7)])

#head of the averaged evaporation dataset of Texas in January 2015 with UTM coordinates
head(JanEvap2015Subset[c(1,2,3,4,7)])

#head of the averaged wind speed dataset of Texas in January 2015 with UTM coordinates
head(JanWind2015Subset[c(1,2,3,4,7)])
```

The data is provided by the National Oceanic and Atmospheric Administration (NOAA) which is a scientific agency within the United Stated Department of Commerce (NOAA, 2018). Environmental data is publicly available on their website and is accessible by an interactive web service. 

## Methods

The two different types of data were analysed with different methods. The time series data of the four stations in Texas were first converted into an `xts` object which provides nice visualizations and makes the handling of them easier. The autocorrelation and partial autocorrelation is computed to see if the data is autocorrelated. The trend of the time series gets extracted with the help of the `stl` function. The trends of the different stations are compared to see differences. A linear model representing the trend is fitted to simplify the comparison. A forecast is made to look on the future development of the climate. Additionally, some basic statistical parameters will be computed. 
The averaged data of all stations were first extended by WGS84 coordinates by converting the given numeric values for Latitude and Longitude. After that the coordinates were transformed into UTM coordinates so that linear units can be used. To compare the data from different years, a kriging interpolation was conducted for every parameter in every given dataset. To be able to do that, a variogram needs to be computed and a suitable model needs to be fitted. For every dataset of the climate elements the same variogram model was fitted to get comparable results. To predict the values of the locations where no measurement was made, a spatial grid is needed. In this case the map of Texas out of the `map` library was used and a regular grid was placed over it. Subsequent, ordinary kriging was performed and the results were visually compared. 

## Results

In this chapter the results will be described, divided into the four investigated climate elements: Temperature, Precipitation, Evaporation and Wind.


### Temperature


#### Time series data

In the four temperature time series of the four stations in Texas from 1950 until today you can see the yearly periodicities as well as the high variability between summer and winter. They all show a general rise of temperature which gets clearer if we look only on the trend.

```{r, echo=FALSE}
#temperature time series of all four stations
plot(tmpMaxAngleton, main="Temperature time series of Angleton in the East of Texas")

plot(tmpMaxBenbrook, main="Temperature time series of Benbrook in the North of Texas")

plot(tmpMaxfortStockton, main="Temperature time series of Fort Stockton in the West of Texas")

plot(tmpMaxMccook, main="Temperature time series of Mccook in the South of Texas")
```

The trend of the temperature time series still shows a lot variability but with the help of the linear model we can see that every time series is increasing.

```{r, echo=FALSE}
#trends of all four temperature time series
plot(trendTmpAngleton)
abline(linTrendAngletonTemp, col="red")
title(main="Trend of temperature time series of Angleton in the East of Texas")

plot(trendTmpBenbrook)
abline(linTrendBenbrookTemp, col="red")
title(main="Trend of temperature time series of Benbrook in the North of Texas")

plot(trendTmpfortStockton)
abline(linTrendfortStocktonTemp, col="red")
title(main = "Trend of temperature time series of Fort Stockton in the West of Texas")

plot(trendTmpMccook)
abline(linTrendMccookTemp, col="red")
title(main="Trend of temperature time series of Mccook in the South of Texas")
```

In the autocorrelation plot up to lag 365 of the time series, you can see the periodicity. In the summer the autocorrelation turns negative because of the deviation from the mean in the other direction.

```{r, echo=FALSE}
#autocorrelation plot of the Angleton temperature time series
plot(acfAngletonTemp, main="Autocorrelation plot of the temperature time series of Angleton, TX")
```

If we look on the forecasts for the next 100 days, we see that for the stations in Benbrook, Fort Stockton and Mccook a further increase is predicted. Only for the station of Angleton the forecast is nearly static but since the series ends in high temperatures a further increase would make no sense.

```{r, echo=FALSE}
#forecasts for the temperature time series
plot(fcastTempAngleton, include=500, main="")
title(main="100 days forecast for the temperature time series of Angleton, TX")

plot(fcastTempBenbrook, include=500, main="")
title(main="100 days forecast for the temperature time series of Benbrook, TX")

plot(fcastTempfortStockton, include=500, main="")
title(main="100 days forecast for the temperature time series of Fort Stockton, TX")

plot(fcastTempMccook, include=500, main="")
title(main="100 days forecast for the temperature time series of Mccook, TX")
```


#### Averaged monthly data

Interestingly, the kriging interpolation results for the averaged maximum temperature data sets of January show no increasing trend. The plots show a clear division of temperature layers starting with the coldest in the north and getting warmer with less distance to the Gulf of Mexico in the south. 

```{r, echo=FALSE}
#kriging interpolation of maximum temperature in Texas in January
spplot(JanTmpMax1955krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"),main="Temperature January 1955")
spplot(JanTmpMax1965krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Temperature January 1965")
spplot(JanTmpMax1975krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Temperature January 1975")
spplot(JanTmpMax1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Temperature January 1985")
spplot(JanTmpMax1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Temperature January 1995")
spplot(JanTmpMax2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Temperature January 2005")
spplot(JanTmpMax2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Temperature January 2015")
```

In addition to that, the comparison of the mean, minimum, maximum and standard deviation shows also no clear trend. The January 2015 was the second coldest after the January 1985 and the January 1965 was the warmest. The difference between 2015 and 1955 displays even a 2.2 degrees cooling. 

```{r, echo=FALSE, results='asis'}
tableDataTemp1 = data.frame(c(JanTmpMax1955Stats$mean,JanTmpMax1965Stats$mean, JanTmpMax1975Stats$mean,JanTmpMax1985Stats$mean,JanTmpMax1995Stats$mean,JanTmpMax2005Stats$mean,JanTmpMax2015Stats$mean, JanTmpMax2015Stats$mean - JanTmpMax1955Stats$mean),c(JanTmpMax1955Stats$min,JanTmpMax1965Stats$min,JanTmpMax1975Stats$min,JanTmpMax1985Stats$min,JanTmpMax1995Stats$min,JanTmpMax2005Stats$min,JanTmpMax2015Stats$min,JanTmpMax2015Stats$min - JanTmpMax1955Stats$min),c(JanTmpMax1955Stats$max,JanTmpMax1965Stats$max,JanTmpMax1975Stats$max,JanTmpMax1985Stats$max,JanTmpMax1995Stats$max,JanTmpMax2005Stats$max,JanTmpMax2015Stats$max, JanTmpMax2015Stats$max - JanTmpMax1955Stats$max), c(JanTmpMax1955Stats$sd,JanTmpMax1965Stats$sd,JanTmpMax1975Stats$sd,JanTmpMax1985Stats$sd,JanTmpMax1995Stats$sd,JanTmpMax2005Stats$sd,JanTmpMax2015Stats$sd, JanTmpMax2015Stats$sd - JanTmpMax1955Stats$sd))
row.names(tableDataTemp1) = c("1955", "1965", "1975", "1985", "1995", "2005", "2015", "Difference 2015 - 1955")
colnames(tableDataTemp1) = c("Mean", "Min", "Max", "Std. Dev.")
knitr::kable(tableDataTemp1, "html", caption="Statistics of averaged Temperature values in January") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

The kriging results for the averaged maximum temperature data sets in August again show no clear trend but are rather very similar to one another. Unlike in January, there are no clear layers, but a very warm region at the border to Mexico. The north and the area at the Gulf of Mexico are colder. A little rise is recognizable when looking at the interpolation classes which are highest in 2015.

```{r, echo=FALSE}
#kriging interpolation of maximum temperature in Texas in August
spplot(AugTmpMax1955krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"),main="Temperature August 1955")
spplot(AugTmpMax1965krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Temperature August 1965")
spplot(AugTmpMax1975krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Temperature August 1975")
spplot(AugTmpMax1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Temperature August 1985")
spplot(AugTmpMax1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Temperature August 1995")
spplot(AugTmpMax2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Temperature August 2005")
spplot(AugTmpMax2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Temperature August 2015")
```

Corresponding to that, the comparison of the statistical values shows only small changes in the mean temperature. The warmest August was 1985 and the coldest was 2005. The difference between 2015 and 1955 shows a rise of nearly one degree in the mean temperature but more than 2.5 degrees in the station with the highest temperatures. 

```{r, echo=FALSE, results='asis'}
tableDataTemp2 = data.frame(c(AugTmpMax1955Stats$mean,AugTmpMax1965Stats$mean, AugTmpMax1975Stats$mean,AugTmpMax1985Stats$mean,AugTmpMax1995Stats$mean,AugTmpMax2005Stats$mean,AugTmpMax2015Stats$mean, AugTmpMax2015Stats$mean - AugTmpMax1955Stats$mean),c(AugTmpMax1955Stats$min,AugTmpMax1965Stats$min,AugTmpMax1975Stats$min,AugTmpMax1985Stats$min,AugTmpMax1995Stats$min,AugTmpMax2005Stats$min,AugTmpMax2015Stats$min,AugTmpMax2015Stats$min - AugTmpMax1955Stats$min),c(AugTmpMax1955Stats$max,AugTmpMax1965Stats$max,AugTmpMax1975Stats$max,AugTmpMax1985Stats$max,AugTmpMax1995Stats$max,AugTmpMax2005Stats$max,AugTmpMax2015Stats$max, AugTmpMax2015Stats$max - AugTmpMax1955Stats$max), c(AugTmpMax1955Stats$sd,AugTmpMax1965Stats$sd,AugTmpMax1975Stats$sd,AugTmpMax1985Stats$sd,AugTmpMax1995Stats$sd,AugTmpMax2005Stats$sd,AugTmpMax2015Stats$sd, AugTmpMax2015Stats$sd - AugTmpMax1955Stats$sd))
row.names(tableDataTemp2) = c("1955", "1965", "1975", "1985", "1995", "2005", "2015", "Difference 2015 - 1955")
colnames(tableDataTemp2) = c("Mean", "Min", "Max", "Std. Dev.")
knitr::kable(tableDataTemp2, "html", caption="Statistics of averaged Temperature values in August") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

Taken both kriging interpolation into account we see that there is no warming in winter but a small constant growth in the summer months.


#### Comparison coast interior

If we compare the slope of the trend of the temperature time series, we see that the rise is not very steep. Nevertheless, we can see an obvious difference between the slope at the coast and in the interior. The slope in the interior is more than twice as steep which could indicate a stronger change of temperature here. A reason for this could be the higher heat storage capacity of the water which weakens the change at the coast.

```{r, echo=FALSE, results='asis'}
coastInteriorTemp = data.frame(c(slopeTemp$slopeAngletonTemp, slopeTemp$slopeMccookTemp, slopeTemp$slopeBenbrookTemp, slopeTemp$slopefortStocktonTemp, (slopeTemp$slopeAngletonTemp+slopeTemp$slopeMccookTemp)/2, (slopeTemp$slopeBenbrookTemp+slopeTemp$slopefortStocktonTemp)/2))
row.names(coastInteriorTemp) = c("Angleton, East (Coast)", "Mccook, South (Coast)", "Benbrook, North (Interior)", "Fort Stockton, West (Interior)", "Mean (Coast)", "Mean(Interior)")
colnames(coastInteriorTemp) = c("Slope of trend")
knitr::kable(coastInteriorTemp, "html", caption="Slope of trend of the temperature time series on the coast and in the interior") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Precipitation


#### Time series data

In the precipitation time series of the four stations you can already see by the different y-axis that the amount of rainfall is very diverse in different parts of Texas. Since there are many peaks in the data it is hard to recognize a trend.

```{r, echo=FALSE}
#Precipitation time series of all four stations
plot(PrcpAngleton, main="Precipitation time series of Angleton in the East of Texas")

plot(PrcpBenbrook, main="Precipitation time series of Benbrook in the North of Texas")

plot(PrcpfortStockton, main="Precipitation time series of Fort Stockton in the West of Texas")

plot(PrcpMccook, main="Precipitation time series of Mccook in the South of Texas")
```

If we only plot the trend of the time series, we see that there is a general rise in precipitation. By comparing these, we see that three of the four stations have an increasing precipitation and only in Mccook the trend is falling a little. Since Mccook had the highest rise in temperature values this could be an explanation for absent precipitation.

```{r, echo=FALSE}
#trends of all four precipitation time series
plot(trendPrcpAngleton)
abline(linTrendAngletonPrcp, col="red")
title(main="Trend of precipitation time series of Angleton in the East of Texas")

plot(trendPrcpBenbrook)
abline(linTrendBenbrookPrcp, col="red")
title(main="Trend of precipitation time series of Benbrook in the North of Texas")

plot(trendPrcpfortStockton)
abline(linTrendfortStocktonPrcp, col="red")
title(main = "Trend of precipitation time series of Fort Stockton in the West of Texas")

plot(trendPrcpMccook)
abline(linTrendMccookPrcp, col="red")
title(main="Trend of precipitation time series of Mccook in the South of Texas")
```

The autocorrelation plot of the precipitation time series of Angleton up to lag 365, shows that there is a significant autocorrelation up to lag 80. This means that the data is autocorrelated but not as strong as the temperature data for example. Unlike the temperature time series, it shows no periodicities since the seasonal component is much weaker here.

```{r, echo=FALSE}
#autocorrelation plot of the Angleton precipitation time series
plot(acfAngletonPrcp, main="Autocorrelation plot of the precipitation time series of Angleton, TX")
```

The forecast for the next 100 days is very constant. Nevertheless, a small increase is predicted for every station. Since all forecasts start at zero precipitation a little increase is not very surprising.

```{r, echo=FALSE}
#forecasts for the precipitation time series
plot(fcastPrcpAngleton, include=500, main="")
title(main="100 days forecast for the precipitation time series of Angleton, TX")

plot(fcastPrcpBenbrook, include=500, main="")
title(main="100 days forecast for the precipitation time series of Benbrook, TX")

plot(fcastPrcpfortStockton, include=500, main="")
title(main="100 days forecast for the precipitation time series of Fort Stockton, TX")

plot(fcastPrcpMccook, include=500, main="")
title(main="100 days forecast for the precipitation time series of Mccook, TX")
```


#### Averaged monthly data

The kriging interpolation of the averaged precipitation data for January supports the findings of the time series analysis and shows an increase in precipitation. It is not the maximum numbers which get higher but the areas with a high amount of rain become larger. The rainfall is always strongest in the regions in the east of Texas, adjacent to the Gulf of Mexico. 

```{r, echo=FALSE}
#kriging interpolation of maximum precipitation in Texas in January
spplot(JanPrcp1955krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"),main="Precipitation January 1955")
spplot(JanPrcp1965krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Precipitation January 1965")
spplot(JanPrcp1975krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Precipitation January 1975")
spplot(JanPrcp1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Precipitation January 1985")
spplot(JanPrcp1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Precipitation January 1995")
spplot(JanPrcp2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Precipitation January 2005")
spplot(JanPrcp2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Precipitation January 2015")
```

In addition to that, the comparison of the mean, minimum, maximum and standard deviation shows a noticeable rise from 1955 to 2015. The mean of the precipitation is twice as high in 2015 compared to ten years before and compared to 1955 the mean has risen by 2.2mm. The standard deviation is most high in 2015 but is still low enough to not be distorted by outliers.

```{r, echo=FALSE, results='asis'}
tableDataPrcp1 = data.frame(c(JanPrcp1955Stats$mean,JanPrcp1965Stats$mean, JanPrcp1975Stats$mean,JanPrcp1985Stats$mean,JanPrcp1995Stats$mean,JanPrcp2005Stats$mean,JanPrcp2015Stats$mean, JanPrcp2015Stats$mean - JanPrcp1955Stats$mean),c(JanPrcp1955Stats$min,JanPrcp1965Stats$min,JanPrcp1975Stats$min,JanPrcp1985Stats$min,JanPrcp1995Stats$min,JanPrcp2005Stats$min,JanPrcp2015Stats$min,JanPrcp2015Stats$min - JanPrcp1955Stats$min),c(JanPrcp1955Stats$max,JanPrcp1965Stats$max,JanPrcp1975Stats$max,JanPrcp1985Stats$max,JanPrcp1995Stats$max,JanPrcp2005Stats$max,JanPrcp2015Stats$max, JanPrcp2015Stats$max - JanPrcp1955Stats$max), c(JanPrcp1955Stats$sd,JanPrcp1965Stats$sd,JanPrcp1975Stats$sd,JanPrcp1985Stats$sd,JanPrcp1995Stats$sd,JanPrcp2005Stats$sd,JanPrcp2015Stats$sd, JanPrcp2015Stats$sd - JanPrcp1955Stats$sd))
row.names(tableDataPrcp1) = c("1955", "1965", "1975", "1985", "1995", "2005", "2015", "Difference 2015 - 1955")
colnames(tableDataPrcp1) = c("Mean", "Min", "Max", "Std. Dev.")
knitr::kable(tableDataPrcp1, "html", caption="Statistics of averaged Precipitation values in January") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

Unlike the kriging results for January, the results for August are not that distinct. There are no big visible changes from 1955 until 2015. Different to January there is not only much precipitation at the coast but also in the north of Texas. The interpolation shows only very little rainfall for most parts of the state and mostly only the areas with the most predicted rainfall are changing.

```{r, echo=FALSE}
#kriging interpolation of maximum precipitation in Texas in August
spplot(AugPrcp1955krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"),main="Precipitation August 1955")
spplot(AugPrcp1965krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Precipitation August 1965")
spplot(AugPrcp1975krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Precipitation August 1975")
spplot(AugPrcp1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Precipitation August 1985")
spplot(AugPrcp1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Precipitation August 1995")
spplot(AugPrcp2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Precipitation August 2005")
spplot(AugPrcp2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Precipitation August 2015")
```

Interestingly, the comparison of the statistical values shows even a small decrease of rainfall in 2015. If you consider that 2005 and 1995 the mean precipitation was always higher than before this could be a sign for an unusual dry August in 2015. The high standard deviation and highest maximum in 2015 are a sign for heavier but more rare rainfall.

```{r, echo=FALSE, results='asis'}
tableDataPrcp2 = data.frame(c(AugPrcp1955Stats$mean,AugPrcp1965Stats$mean, AugPrcp1975Stats$mean,AugPrcp1985Stats$mean,AugPrcp1995Stats$mean,AugPrcp2005Stats$mean,AugPrcp2015Stats$mean, AugPrcp2015Stats$mean - AugPrcp1955Stats$mean),c(AugPrcp1955Stats$min,AugPrcp1965Stats$min,AugPrcp1975Stats$min,AugPrcp1985Stats$min,AugPrcp1995Stats$min,AugPrcp2005Stats$min,AugPrcp2015Stats$min,AugPrcp2015Stats$min - AugPrcp1955Stats$min),c(AugPrcp1955Stats$max,AugPrcp1965Stats$max,AugPrcp1975Stats$max,AugPrcp1985Stats$max,AugPrcp1995Stats$max,AugPrcp2005Stats$max,AugPrcp2015Stats$max, AugPrcp2015Stats$max - AugPrcp1955Stats$max), c(AugPrcp1955Stats$sd,AugPrcp1965Stats$sd,AugPrcp1975Stats$sd,AugPrcp1985Stats$sd,AugPrcp1995Stats$sd,AugPrcp2005Stats$sd,AugPrcp2015Stats$sd, AugPrcp2015Stats$sd - AugPrcp1955Stats$sd))
row.names(tableDataPrcp2) = c("1955", "1965", "1975", "1985", "1995", "2005", "2015", "Difference 2015 - 1955")
colnames(tableDataPrcp2) = c("Mean", "Min", "Max", "Std. Dev.")
knitr::kable(tableDataPrcp2, "html", caption="Statistics of averaged Precipitation values in August") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

If you take both kriging results into account, you could speak of a small rise in precipitation even if the interpolation of August is not completely distinct.  


#### Comparison coast interior

Matching the results of the kriging interpolation the comparison of the slope of the trend for the precipitation time series shows that the increase at the coast is three times higher than in the interior. Even if the slope in Mccook is decreasing, the steep slope in Angleton sets that off. Different to the temperature time series the changes are now higher at the coast.

```{r, echo=FALSE, results='asis'}
coastInteriorPrcp = data.frame(c(slopePrcp$slopeAngletonPrcp, slopePrcp$slopeMccookPrcp, slopePrcp$slopeBenbrookPrcp, slopePrcp$slopefortStocktonPrcp, (slopePrcp$slopeAngletonPrcp+slopePrcp$slopeMccookPrcp)/2, (slopePrcp$slopeBenbrookPrcp+slopePrcp$slopefortStocktonPrcp)/2))
row.names(coastInteriorPrcp) = c("Angleton, East (Coast)", "Mccook, South (Coast)", "Benbrook, North (Interior)", "Fort Stockton, West (Interior)", "Mean (Coast)", "Mean(Interior)")
colnames(coastInteriorPrcp) = c("Slope of trend")
knitr::kable(coastInteriorPrcp, "html", caption="Slope of trend of the Precipitation time series on the coast and in the interior") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Evaporation


#### Time series data

The evaporation time series of the four stations in Texas do not cover the whole extent from 1950 until today. The time series of Angleton and Fort Stockton both start in the early 1950's but end only ten years later in the early 1960's. The time series of Mccook and Benbrook start in 1963 and 1979 but both reach until today. To be consistent, the measurements were still used in the analysis, but the results are not as meaningful as for the other measurements. Like in the temperature time series we can see yearly periodicities. In none of the time series a clear trend is visible but for the two stations with the longest series the peaks get higher with time.

```{r, echo=FALSE}
#Evaporation time series of all four stations
plot(EvapAngleton, main="Evaporation time series of Angleton in the East of Texas")

plot(EvapBenbrook, main="Evaporation time series of Benbrook in the North of Texas")

plot(EvapfortStockton, main="Evaporation time series of Fort Stockton in the West of Texas")

plot(EvapMccook, main="Evaporation time series of Mccook in the South of Texas")
```

Different from the assumption that the evaporation should be increasing due to higher temperatures, the trend of three of the four stations is decreasing. Even if we leave out the stations with too little data, the time series of the longest period also shows a decreasing trend. According to research the increased cloud coverage and aerosol concentration leads to a decreased evaporation (Roderick and Farquhar, 2002). Nevertheless, the evaporation on sunny days seems to be stronger than former because of higher peaks at the end of the time series.

```{r, echo=FALSE}
#trends of all four Evaporation time series
plot(trendEvapAngleton)
abline(linTrendAngletonEvap, col="red")
title(main="Trend of Evaporation time series of Angleton in the East of Texas")

plot(trendEvapBenbrook)
abline(linTrendBenbrookEvap, col="red")
title(main="Trend of Evaporation time series of Benbrook in the North of Texas")

plot(trendEvapfortStockton)
abline(linTrendfortStocktonEvap, col="red")
title(main = "Trend of Evaporation time series of Fort Stockton in the West of Texas")

plot(trendEvapMccook)
abline(linTrendMccookEvap, col="red")
title(main="Trend of Evaporation time series of Mccook in the South of Texas")
```

The autocorrelation plot for Angleton shows the seasonal periodicities of the evaporation. In the winter there is only very little evaporation compared to the summer.

```{r, echo=FALSE}
#autocorrelation plot of the Angleton Evaporation time series
plot(acfAngletonEvap, main="Autocorrelation plot of the Evaporation time series of Angleton, TX")
```

According to the decreasing trend, also most of the forecasts expect a further decrease of evaporation. Interestingly, the forecast for Mccook is strongly rising in the first days although it has a decreasing trend. Overall the forecasts are rather static with only small changes.

```{r, echo=FALSE}
#forecasts for the Evaporation time series
plot(fcastEvapAngleton, include=500, main="")
title(main="100 days forecast for the Evaporation time series of Angleton, TX")

plot(fcastEvapBenbrook, include=500, main="")
title(main="100 days forecast for the Evaporation time series of Benbrook, TX")

plot(fcastEvapfortStockton, include=500, main="")
title(main="100 days forecast for the Evaporation time series of Fort Stockton, TX")

plot(fcastEvapMccook, include=500, main="")
title(main="100 days forecast for the Evaporation time series of Mccook, TX")
```


#### Averaged monthly data

The kriging interpolation for January is not showing a clear trend in the one or the other direction. Matching to the kriging results of temperature we see that the evaporation is strongest in the regions where temperature is high. In this case this is at the border to Mexico in the southwest of Texas.

```{r, echo=FALSE}
#kriging interpolation of maximum Evaporation in Texas in January
spplot(JanEvap1955krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"),main="Evaporation January 1955")
spplot(JanEvap1965krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Evaporation January 1965")
spplot(JanEvap1975krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Evaporation January 1975")
spplot(JanEvap1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Evaporation January 1985")
spplot(JanEvap1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Evaporation January 1995")
spplot(JanEvap2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Evaporation January 2005")
spplot(JanEvap2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Evaporation January 2015")
```

According to the previous findings the comparison of the basic statistics shows that there is no clear trend for the evaporation, but it rather seems to decrease than to increase. Another interesting thing is that the maximum evaporation and the standard deviation are highest in 2015 indicating that on sunny days the evaporation is stronger nowadays and the variations become higher.

```{r, echo=FALSE, results='asis'}
tableDataEvap1 = data.frame(c(JanEvap1955Stats$mean,JanEvap1965Stats$mean, JanEvap1975Stats$mean,JanEvap1985Stats$mean,JanEvap1995Stats$mean,JanEvap2005Stats$mean,JanEvap2015Stats$mean, JanEvap2015Stats$mean - JanEvap1955Stats$mean),c(JanEvap1955Stats$min,JanEvap1965Stats$min,JanEvap1975Stats$min,JanEvap1985Stats$min,JanEvap1995Stats$min,JanEvap2005Stats$min,JanEvap2015Stats$min,JanEvap2015Stats$min - JanEvap1955Stats$min),c(JanEvap1955Stats$max,JanEvap1965Stats$max,JanEvap1975Stats$max,JanEvap1985Stats$max,JanEvap1995Stats$max,JanEvap2005Stats$max,JanEvap2015Stats$max, JanEvap2015Stats$max - JanEvap1955Stats$max), c(JanEvap1955Stats$sd,JanEvap1965Stats$sd,JanEvap1975Stats$sd,JanEvap1985Stats$sd,JanEvap1995Stats$sd,JanEvap2005Stats$sd,JanEvap2015Stats$sd, JanEvap2015Stats$sd - JanEvap1955Stats$sd))
row.names(tableDataEvap1) = c("1955", "1965", "1975", "1985", "1995", "2005", "2015", "Difference 2015 - 1955")
colnames(tableDataEvap1) = c("Mean", "Min", "Max", "Std. Dev.")
knitr::kable(tableDataEvap1, "html", caption="Statistics of averaged Evaporation values in January") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

The kriging interpolation results for August show no clear trend as well. Interestingly, the area with the strongest evaporation is now the northwest at the border to New Mexico. As expected the evaporation values are higher than in January since the temperatures are also a lot higher in August.

```{r, echo=FALSE}
#kriging interpolation of maximum Evaporation in Texas in August
spplot(AugEvap1955krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"),main="Evaporation August 1955")
spplot(AugEvap1965krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Evaporation August 1965")
spplot(AugEvap1975krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Evaporation August 1975")
spplot(AugEvap1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Evaporation August 1985")
spplot(AugEvap1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Evaporation August 1995")
spplot(AugEvap2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Evaporation August 2005")
spplot(AugEvap2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Evaporation August 2015")
```

The basic statistical values in August are similar to the ones from January. Beside the values in 2015 it looks more like a decrease than an increase. Again, the maximum evaporation value and highest standard deviation is from 2015. This indicates a stronger but more rare evaporation.

```{r, echo=FALSE, results='asis'}
tableDataEvap2 = data.frame(c(AugEvap1955Stats$mean,AugEvap1965Stats$mean, AugEvap1975Stats$mean,AugEvap1985Stats$mean,AugEvap1995Stats$mean,AugEvap2005Stats$mean,AugEvap2015Stats$mean, AugEvap2015Stats$mean - AugEvap1955Stats$mean),c(AugEvap1955Stats$min,AugEvap1965Stats$min,AugEvap1975Stats$min,AugEvap1985Stats$min,AugEvap1995Stats$min,AugEvap2005Stats$min,AugEvap2015Stats$min,AugEvap2015Stats$min - AugEvap1955Stats$min),c(AugEvap1955Stats$max,AugEvap1965Stats$max,AugEvap1975Stats$max,AugEvap1985Stats$max,AugEvap1995Stats$max,AugEvap2005Stats$max,AugEvap2015Stats$max, AugEvap2015Stats$max - AugEvap1955Stats$max), c(AugEvap1955Stats$sd,AugEvap1965Stats$sd,AugEvap1975Stats$sd,AugEvap1985Stats$sd,AugEvap1995Stats$sd,AugEvap2005Stats$sd,AugEvap2015Stats$sd, AugEvap2015Stats$sd - AugEvap1955Stats$sd))
row.names(tableDataEvap2) = c("1955", "1965", "1975", "1985", "1995", "2005", "2015", "Difference 2015 - 1955")
colnames(tableDataEvap2) = c("Mean", "Min", "Max", "Std. Dev.")
knitr::kable(tableDataEvap2, "html", caption="Statistics of averaged Evaporation values in August") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

If we look at both kriging results it seems like the evaporation is not changing very strongly with the climate change. According to the findings of Roderick and Farquhar there is a weak decreasing trend.


#### Comparison coast interior

The comparison of the slope of trends also shows this decrease in evaporation but must be considered carefully since the time series not cover the same temporal extent. The comparison indicates that the decrease in the interior is higher than at the coast. This stands in contrast to the higher increase of temperature in the interior. 

```{r, echo=FALSE, results='asis'}
coastInteriorEvap = data.frame(c(slopeEvap$slopeAngletonEvap, slopeEvap$slopeMccookEvap, slopeEvap$slopeBenbrookEvap, slopeEvap$slopefortStocktonEvap, (slopeEvap$slopeAngletonEvap+slopeEvap$slopeMccookEvap)/2, (slopeEvap$slopeBenbrookEvap+slopeEvap$slopefortStocktonEvap)/2))
row.names(coastInteriorEvap) = c("Angleton, East (Coast)", "Mccook, South (Coast)", "Benbrook, North (Interior)", "Fort Stockton, West (Interior)", "Mean (Coast)", "Mean(Interior)")
colnames(coastInteriorEvap) = c("Slope of trend")
knitr::kable(coastInteriorEvap, "html", caption="Slope of trend of the Evaporation time series on the coast and in the interior") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Wind


#### Time series data

Like the evaporation time series, the total wind movement time series do not cover the full extent from 1950 until today. Again, the stations of Benbrook (1979 - today) and Mccook (1963 - 2003) cover longer periods than the stations in Angleton (1953 - 1963) and Fort Stockton (1950 - 1960). The total wind movement represents the summed wind speeds for one day. We can see yearly periodicities of the time series but no clear trend.

```{r, echo=FALSE}
#Total Wind Movement time series of all four stations
plot(WindAngleton, main="Total Wind Movement time series of Angleton in the East of Texas")

plot(WindBenbrook, main="Total Wind Movement time series of Benbrook in the North of Texas")

plot(WindfortStockton, main="Total Wind Movement time series of Fort Stockton in the West of Texas")

plot(WindMccook, main="Total Wind Movement time series of Mccook in the South of Texas")
```

If we compare only the trends of the time series, we see a very ambiguous pattern. The time series of Angleton and Benbrook have a weak increasing trend and the time series of Fort Stockton and Mccook have a strong decreasing trend. If we only consider the time series with the longer observation periods we have one with an increasing trend (Benbrook) and one with a decreasing trend (Mccook).

```{r, echo=FALSE}
#trends of all four Total Wind Movement time series
plot(trendWindAngleton)
abline(linTrendAngletonWind, col="red")
title(main="Trend of Total Wind Movement time series of Angleton in the East of Texas", cex.main=1)

plot(trendWindBenbrook)
abline(linTrendBenbrookWind, col="red")
title(main="Trend of Total Wind Movement time series of Benbrook in the North of Texas", cex.main=1)

plot(trendWindfortStockton)
abline(linTrendfortStocktonWind, col="red")
title(main = "Trend of Total Wind Movement time series of Fort Stockton in the West of Texas", cex.main=1)

plot(trendWindMccook)
abline(linTrendMccookWind, col="red")
title(main="Trend of Total Wind Movement time series of Mccook in the South of Texas", cex.main=1)
```

The autocorrelation plot shows the same yearly periodicity as the temperature and the evaporation time series since the wind is stronger in the winter.

```{r, echo=FALSE}
#autocorrelation plot of the Angleton Total Wind Movement time series
plot(acfAngletonWind, main="")
title(main="Autocorrelation plot of the Total Wind Movement time series of Angleton, TX", cex.main=1)
```

According to the trends of the time series the forecast for Angleton and Mccook are also increasing respectively decreasing. Interestingly, the forecasts for the other two stations are different than their trends. For Benbrook the forecast predicts a strong decrease in opposite to the increasing trend. An explanation for this could be that the last measurement is very high (over 200) and a decrease is most likely to occur. For Fort Stockton an increase is predicted although the trend is strongly decreasing.

```{r, echo=FALSE}
#forecasts for the Total Wind Movement time series
plot(fcastWindAngleton, include=500, main="")
title(main="100 days forecast for the Total Wind Movement time series of Angleton, TX", cex.main=1)

plot(fcastWindBenbrook, include=500, main="")
title(main="100 days forecast for the Total Wind Movement time series of Benbrook, TX", cex.main=1)

plot(fcastWindfortStockton, include=500, main="")
title(main="100 days forecast for the Total Wind Movement time series of Fort Stockton, TX", cex.main=1)

plot(fcastWindMccook, include=500, main="")
title(main="100 days forecast for the Total Wind Movement time series of Mccook, TX", cex.main=1)
```


#### Averaged monthly data

For the kriging interpolation the averaged daily average wind speeds are used. Since these measurements are only observed since 1980 the kriging includes only the years 1985, 1995, 2005 and 2015. The kriging results for January show no clear trend at first glance. The visualizations are a bit confusing since the results for 2015 and 2005 have a very high maximum wind speed class which only applies for very small regions in the west of Texas. 

```{r, echo=FALSE}
#kriging interpolation of maximum Wind Speed in Texas in January
spplot(JanWind1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Wind Speed January 1985")
spplot(JanWind1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Wind Speed January 1995")
spplot(JanWind2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Wind Speed January 2005")
spplot(JanWind2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Wind Speed January 2015")
```

By comparing the statistical values, we see a distinct decreasing trend in wind speed. The average wind speed of 1985 is about 0.7 m/h faster compared to the average wind speed of 2015. Only the variability gets stronger which is represented by the highest standard deviation in 2015. We have on average a slower wind speed but more days with an unusual high wind speed.

```{r, echo=FALSE, results='asis'}
tableDataWind1 = data.frame(c(JanWind1985Stats$mean,JanWind1995Stats$mean,JanWind2005Stats$mean,JanWind2015Stats$mean, JanWind2015Stats$mean - JanWind1985Stats$mean),c(JanWind1985Stats$min,JanWind1995Stats$min,JanWind2005Stats$min,JanWind2015Stats$min,JanWind2015Stats$min - JanWind1985Stats$min),c(JanWind1985Stats$max,JanWind1995Stats$max,JanWind2005Stats$max,JanWind2015Stats$max, JanWind2015Stats$max - JanWind1985Stats$max), c(JanWind1985Stats$sd,JanWind1995Stats$sd,JanWind2005Stats$sd,JanWind2015Stats$sd, JanWind2015Stats$sd - JanWind1985Stats$sd))
row.names(tableDataWind1) = c("1985", "1995", "2005", "2015", "Difference 2015 - 1985")
colnames(tableDataWind1) = c("Mean", "Min", "Max", "Std. Dev.")
knitr::kable(tableDataWind1, "html", caption="Statistics of averaged Wind Speed values in January") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

In August we have the same confusing visualizations of 2015 and 2005 like in January since some high measurements disturb the interpolation. By this it is again hard to recognize a trend. Like in January the highest wind speeds appear in the west and in the north of Texas. As expected the wind speeds in august are much slower compared to January.

```{r, echo=FALSE}
#kriging interpolation of maximum Wind Speed in Texas in August
spplot(AugWind1985krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Wind Speed August 1985")
spplot(AugWind1995krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Wind Speed August 1995")
spplot(AugWind2005krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Wind Speed August 2005")
spplot(AugWind2015krige1[1], sp.layout=list(texasMap, first=FALSE, lwd=2, col="white"), main="Wind Speed August 2015")
```

Unlike in January the comparison of the basic statistical values shows no clear trend. If we would leave out 2015 there is again a decreasing trend. 2015 the data has the largest variability regarding the difference between the minimal and maximal value. Different from before the standard deviation is not highest in 2015.

```{r, echo=FALSE, results='asis'}
tableDataWind2 = data.frame(c(AugWind1985Stats$mean,AugWind1995Stats$mean,AugWind2005Stats$mean,AugWind2015Stats$mean, AugWind2015Stats$mean - AugWind1985Stats$mean),c(AugWind1985Stats$min,AugWind1995Stats$min,AugWind2005Stats$min,AugWind2015Stats$min,AugWind2015Stats$min - AugWind1985Stats$min),c(AugWind1985Stats$max,AugWind1995Stats$max,AugWind2005Stats$max,AugWind2015Stats$max, AugWind2015Stats$max - AugWind1985Stats$max), c(AugWind1985Stats$sd,AugWind1995Stats$sd,AugWind2005Stats$sd,AugWind2015Stats$sd, AugWind2015Stats$sd - AugWind1985Stats$sd))
row.names(tableDataWind2) = c("1985", "1995", "2005", "2015", "Difference 2015 - 1985")
colnames(tableDataWind2) = c("Mean", "Min", "Max", "Std. Dev.")
knitr::kable(tableDataWind2, "html", caption="Statistics of averaged Wind Speed values in August") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

Summarized the wind speed seems to decrease with the climate change. Since there was a clear decreasing trend in the data sets for January and the differences in the data sets for August were only very small, this indicates an overall decreasing trend.


#### Comparison coast interior

For the first time the comparison of the slope of the time series trends shows no difference between coast and interior. This result must be considered carefully since the time series have not the same temporal extent. Nevertheless, it again shows a decreasing overall trend of the wind movement. 

```{r, echo=FALSE, results='asis'}
coastInteriorWind = data.frame(c(slopeWind$slopeAngletonWind, slopeWind$slopeMccookWind, slopeWind$slopeBenbrookWind, slopeWind$slopefortStocktonWind, (slopeWind$slopeAngletonWind+slopeWind$slopeMccookWind)/2, (slopeWind$slopeBenbrookWind+slopeWind$slopefortStocktonWind)/2))
row.names(coastInteriorWind) = c("Angleton, East (Coast)", "Mccook, South (Coast)", "Benbrook, North (Interior)", "Fort Stockton, West (Interior)", "Mean (Coast)", "Mean(Interior)")
colnames(coastInteriorWind) = c("Slope of trend")
knitr::kable(coastInteriorWind, "html", caption="Slope of trend of the Total Wind Movement time series on the coast and in the interior") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Discussion

The results of the analysis show that there is a remarkable climate change in Texas over the last 70 years. The temperature has risen by about 1 degree in August and is predicted to further increase. The precipitation shows a small increase by about 2 mm in January and is also predicted to further increase. The evaporation is decreasing even if the results are not completely distinct. The wind speed and total wind movement are also decreasing.
There are differences in the magnitude of the climate change at the coast and in the interior. At the coast the increase of temperature is stronger as well as the decrease of evaporation while the increase of precipitation is stronger in the interior. The decrease of wind speed and wind movement shows no differences between coast and interior.
Since most times the maximum values were most high in 2015 this could indicate more extreme climate conditions which could be a starting point for further research.
Unfortunately, the evaporation and total wind movement time series do not cover the full temporal extent from 1950 until today but sometimes only ten years of data. This could influence the results since they were analysed in the same way as the other time series. With a more complete data set the results would be more meaningful. According to that it would be better to take more than four stations into account to get more credible results. The same applies for the kriging interpolation. With more data from different months or also additional years the results could be different and would be more meaningful. Since the used data already has a size of about 100 megabyte it would have been too computationally intensive to use more. The chosen data seemed to be the best fitting according to the research question.

## Conclusion

There is a remarkable climate change in Texas regarding the climate elements temperature, precipitation, evaporation and wind speed. By this the work provides an additional evidence of the climate change and adds to the multitude of research in this field. 
Further research should use more and different data for this analysis to prove these results and create more meaningful outcomes. Furthermore, other methods for the analysis could be used like dynamic time warping for example which provides algorithms to compare time series (Dynamic Time Warp Project, 2018).

## References

DallasNews, 2007. "Texas No. 1 producer of greenhouse gases". Retrieved March 13, 2018 from: <https://web.archive.org/web/20080919052620/http://www.dallasnews.com/sharedcontent/dws/news/texassouthwest/stories/060307dnnatemissions.3c1df3a.html>

Dynamic Time Warp Project, 2018. "Dynamic Time Warp project". Retrieved March 15, 2018 from: <http://dtw.r-forge.r-project.org/>

IPCC, 2018. "Intergovernmental Panel On Climate Change". Retrieved March 13, 2018 from: <http://www.ipcc.ch/>

IPCC Working Group 1, 2007. "Climate Change 2007: The Physical Science Basis". Contribution of Working Group I to the Fourth Assessment Report of the Intergovernmental Panel on Climate Change

IPCC Working Group 2, 2007. "Climate Change 2007: Impacts, Adaptation and Vulnerability". Working Group II Contribution to the Intergovernmental Panel on Climate Change Fourth Assessment Report

Jouzel, J. et al. 2007. "Orbital and Millennial Antarctic Climate Variability over the Past 800,000 Years". Science 317 (5839), 793-796. DOI: 10.1126/science.1141038

Keeling, C. D., 1960. "The concentration and isotopic abundances of carbon dioxide in the atmosphere". Tellus Volume 12, Number 2

NOAA, 2018. "Climate Data Online Search". Retrieved March 13, 2018 from: <https://www.ncdc.noaa.gov/cdo-web/search>

Roderick, M. L. and Farquhar, G. D., 2002. "The Cause of Decreased Pan Evaporation over the Past 50 Years". Science 298 (5597), 1410-1411. DOI: 10.1126/science.1075390-a

Schmandt, J., North, G. R. and Clarkson, J., 2011. "The impact of global warming on Texas". 2nd Edition. Austin, TX. University of Texas Press

Texas Almanac, 2018. "Texas State Historical Association, Facts". Retrieved March 13, 2018 from: <http://texasalmanac.com/topics/facts-profile>